### 2019-04-02 Publication/manuscript.tex
Create manuscript.tex

\documentclass[a4paper]{article}
\usepackage{authblk}
\usepackage[backend=bibtex]{biblatex}
\usepackage{hyperref}
\usepackage{txfonts}
\usepackage{titling}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[a4paper, margin={2.5cm, 1.7cm}]{geometry}
\usepackage{mwe}
\usepackage{filecontents}
\usepackage{changes}
\usepackage{lipsum}% <- For dummy text
\definechangesauthor[name={Heide}, color=blue]{Heide}
\setremarkmarkup{(#2)}

\bibliography{database.bib}

\begin{document}
\title{Fast processing of Jungfrau detector data}
\author{Jonas Schenke$^{1,2}$, 
Florian Warg$^{1,2}$, 
Anna Bergamaschi$^3$,
Martin Br\"uckner$^3$,\\
Michael Bussmann$^1$,
Carlos Lopez-Cuenza$^3$,
Aldo Mozzanica$^3$,\\
Sophie Redford$^3$,
Bernd Schmitt$^3$,
Heide Mei{\ss}ner}

\affil[1]{HZDR}
\affil[2]{TU Dresden}
\affil[3]{PSI}

\date{}

\renewcommand\Affilfont{\itshape}

\maketitle
{\bf Abstract}\\
...text...\\

{\bf Keywords:} Photon pixel detector, fast data processing, GPU programming, Alpaka...

\section{Introduction}

Increasing data rates during FEL experiments require dedicated detectors as well as advanced methods for fast data processing\\

Jungfrau detector (\cite{jungfrau1}, \cite{jungfrau2}, \cite{jungfrau3}): pixel detector with gain switching scheme for large range of photon rates (single pixel to photon bunches)\\

online data conversion: calculate energy and number of photons from detector response for each pixel using continuously updated pedestal maps\\

Find clusters\\

Hardware-independent computation\\

Different numbers of modules, parallel processing\\

GPU / Alpaka: \cite{Matthes17}: portable, parallel and scalable code\\

Related work\\

...\\

In the following, we describe ....\\

\section{Methods}
\subsection{Abilities and applications of the Jungfrau Detector (PSI)}

...

\begin{figure}[h!]
\centering
\includegraphics[width=0.30\textwidth]{jungfraudetector.jpg}
\caption{Jungfrau detector}
\label{fig:jfdetector}
\end{figure}

\subsection{Data processing algorithm (PSI)}
conversion from detector data to energy and number of photons\\

 summation of frames\\

 clustering (reference?)\\

\subsection{Alpaka implementation of fast data processing (Jonas, Florian)}

\subsection{Benchmark tests (Jonas, Florian)}
Design, objectives, and evaluation of tests

\section{Results}
\subsection{Achieved improvements (Jonas)}
Results of tests of software parts on various computing hardware using suitable Alpaka backends\\

Where are the bottlenecks\\

Available capacity on GPUs\\

Best system

\subsection{Experiences from practical application of improved code (PSI)}
Application results

\section{Conclusions and Outlook}

Is presented method applicable / useful for other detectors, e.g. AGIPD? \\

calculation on FPGAs in the future

\section{Acknowledges}
Thanks to Alpaka developers.....\\

This project has received funding from the European Union's Horizon 2020 research and innovation programme under grant agreement No 654220 (EUCALL).

\newpage

\begin{sloppypar}
\printbibliography
\end{sloppypar}
\end{document}

### 2019-04-03 Publication/manuscript.tex
Few notes to remember and JF ASIC ref.

@@
 
-Jungfrau detector (\cite{jungfrau1}, \cite{jungfrau2}, \cite{jungfrau3}): pixel detector with gain switching scheme for large range of photon rates (single pixel to photon bunches)\\
+Jungfrau detector (\cite{jungfrau1}, \cite{jungfrau2}, \cite{jungfrau3}, \cite{Mozzanica_2016}): pixel detector with gain switching scheme for large range of photon rates (single pixel to photon bunches)\\
 
-online data conversion: calculate energy and number of photons from detector response for each pixel using continuously updated pedestal maps\\
+charge integrating, meaning that detector output must be corrected before the number of photons per pixel per frame is retrieved.\\
+
+online data conversion: calculate energy and number of photons from detector response for each pixel using continuously updated pedestal maps and gain maps calculated in dedicated lab-based calibration procedure~\cite{jungfrau2}\\
 
@@
 \subsection{Data processing algorithm (PSI)}
-conversion from detector data to energy and number of photons\\
+conversion from detector data to energy and number of photons including pixel mask, pedestal calculation, pedestal tracking, pedestal correction, gain correction, conversion to number of photons (in the case of monochromatic incident radiation)\\
 

### 2019-04-17 include/Filecache.hpp
Templated everything.

@@ -4,63 +4,74 @@
 
 #include <fstream>
 #include <memory>
+#include <sys/stat.h>
+#include <sys/types.h>
+#include <unistd.h>
 
-
-class Filecache {
+template <typename Config> class Filecache {
 private:
-  std::unique_ptr<char, std::function<void(char*)>> buffer;
+    std::unique_ptr<char, std::function<void(char*)>> buffer;
     char* bufferPointer;
     const std::size_t sizeBytes;
-    auto getFileSize(const std::string path) const -> off_t;
+    auto getFileSize(const std::string path) const -> off_t
+    {
+        struct stat fileStat;
+        stat(path.c_str(), &fileStat);
+        return fileStat.st_size;
+    }
 
 public:
-    Filecache(std::size_t size);
+    Filecache(std::size_t size)
+        : buffer(
+              [size]() {
+                  void* p(malloc(size));
+                  return reinterpret_cast<char*>(p);
+              }(),
+              free),
+          bufferPointer(buffer.get()),
+          sizeBytes(size)
+    {
+    }
+  
     template <typename TData, typename TAlpaka>
     auto loadMaps(const std::string& path, bool header = false)
-        -> FramePackage<TData, TAlpaka>;
+        -> typename Config::template FramePackage<TData, TAlpaka>
+    {
+        // allocate space
+        auto fileSize = getFileSize(path);
+        std::size_t numFrames = fileSize / sizeof(TData);
+
+        if (fileSize + bufferPointer >= buffer.get() + sizeBytes) {
+            std::cerr << "Error: Not enough memory allocated!\n";
+            exit(EXIT_FAILURE);
+        }
+
+        // load file content
+        std::ifstream file;
+        file.open(path, std::ios::in | std::ios::binary);
+        if (!file.is_open()) {
+            std::cerr << "Error: Couldn't open file " << path << "!\n";
+            exit(EXIT_FAILURE);
+        }
+
+        file.read(bufferPointer, fileSize);
+        file.close();
+
+        typename Config::template FramePackage<TData, TAlpaka> maps(numFrames);
+
+        CpuSyncQueue streamBuf = alpakaGetHost<TAlpaka>();
+
+        TData* dataBuf = reinterpret_cast<TData*>(bufferPointer);
+
+        // copy data into alpaca memory
+        alpakaCopy(streamBuf,
[…]

### 2019-04-17 include/jungfrau-photoncounter/Alpakaconfig.hpp

@@ -1,6 +1,6 @@
 /**
  * \file
- * Copyright 2018 Sebastian Benner, Jonas Schenke
+ * Copyright 2019 Sebastian Benner, Jonas Schenke
  *
  * This file is part of jungfrau-photoncounter.
  *
@@ -22,7 +22,6 @@
 
 #pragma once
 
-#include "Config.hpp"
 #include "kernel/Iterator.hpp"
 #include <alpaka/alpaka.hpp>
 
@@ -69,6 +68,8 @@ struct CpuTbbBlocks {
 };
 #endif
 
+
+/*
 template <typename THost, typename TAcc, typename TQueue> struct AccTraits {
     using Host = THost;
     using Acc = TAcc;
@@ -85,16 +86,30 @@ template <typename THost, typename TAcc, typename TQueue> struct AccTraits {
     template <typename T>
     using HostView = alpaka::mem::view::ViewSubView<DevHost, T, Dim, Size>;
 };
+*/
+
 
 #ifdef ALPAKA_ACC_CPU_B_OMP2_T_SEQ_ENABLED
 //#############################################################################
 //! OpenMP 2 Blocks defines
 //!
 //! Defines Host, Device, etc. for the OpenMP 2 Blocks accelerator.
-struct CpuOmp2Blocks
-    : public AccTraits<alpaka::acc::AccCpuOmp2Blocks<Dim, Size>,
-                       alpaka::acc::AccCpuOmp2Blocks<Dim, Size>,
-                       alpaka::queue::QueueCpuSync> {
+template <std::size_t MAPSIZE> struct CpuOmp2Blocks {
+    using Host = alpaka::acc::AccCpuOmp2Blocks<Dim, Size>;
+    using Acc = alpaka::acc::AccCpuOmp2Blocks<Dim, Size>;
+    using Queue = alpaka::queue::QueueCpuSync;
+    using DevHost = alpaka::dev::Dev<Host>;
+    using DevAcc = alpaka::dev::Dev<Acc>;
+    using PltfHost = alpaka::pltf::Pltf<DevHost>;
+    using PltfAcc = alpaka::pltf::Pltf<DevAcc>;
+    using Event = alpaka::event::Event<Queue>;
+    template <typename T>
+    using AccBuf = alpaka::mem::buf::Buf<DevAcc, T, Dim, Size>;
+    template <typename T>
+    using HostBuf = alpaka::mem::buf::Buf<DevHost, T, Dim, Size>;
+    template <typename T>
[…]

### 2019-04-17 include/jungfrau-photoncounter/Config.hpp

@@ -5,130 +5,142 @@
 #include <fstream>
 #include <iostream>
 
-#include "CheapArray.hpp"
 #include "AlpakaHelper.hpp"
+#include "CheapArray.hpp"
 
-// general settings
-constexpr std::size_t FRAMESPERSTAGE_G0 = 1000;
-constexpr std::size_t FRAMESPERSTAGE_G1 = 0;
-constexpr std::size_t FRAMESPERSTAGE_G2 = 0;
+template <std::size_t TFramesPerStageG0,
+          std::size_t TFramesPerStageG1,
+          std::size_t TFramesPerStageG2,
+          std::size_t TDimX,
+          std::size_t TDimY,
+          std::size_t TSumFrames,
+          std::size_t TDevFrames,
+          std::size_t TMovingStatWindowSize,
+          std::size_t TClusterSize,
+          std::size_t TC>
+struct DetectorConfig {
+    static constexpr std::size_t FRAMESPERSTAGE_G0 = TFramesPerStageG0;
+    static constexpr std::size_t FRAMESPERSTAGE_G1 = TFramesPerStageG1;
+    static constexpr std::size_t FRAMESPERSTAGE_G2 = TFramesPerStageG2;
+    static constexpr std::size_t DIMX = TDimX;
+    static constexpr std::size_t DIMY = TDimY;
+    static constexpr std::size_t SUM_FRAMES = TSumFrames;
+    static constexpr std::size_t DEV_FRAMES = TDevFrames;
+    static constexpr std::size_t MOVING_STAT_WINDOW_SIZE = TMovingStatWindowSize;
+    static constexpr std::size_t CLUSTER_SIZE = TClusterSize;
+    static constexpr std::size_t C = TC;
 
-constexpr std::size_t FRAME_HEADER_SIZE = 16;
-constexpr std::size_t DIMX = 400;
-constexpr std::size_t DIMY = 400;
-constexpr std::size_t SUM_FRAMES = 10;
-constexpr std::size_t DEV_FRAMES = 300;
-constexpr std::size_t PEDEMAPS = 3;
-constexpr std::size_t MOVING_STAT_WINDOW_SIZE = 100;
-constexpr std::size_t GAINMAPS = 3;
-constexpr float BEAMCONST = 6.2;
-constexpr float PHOTONCONST = (1. / 12.4);
-constexpr int CLUSTER_SIZE = 3;
-constexpr int C = 5;
+    // general settings
[…]

### 2019-04-17 include/jungfrau-photoncounter/Debug.hpp

@@ -7,15 +7,15 @@
 
 #include "Config.hpp"
 
-template <typename TAlpaka>
-void saveClusters(std::string path,
-                  ClusterArray<TAlpaka>& clusters)
+template <typename TConfig, typename TAlpaka>
+void saveClusters(std::string path, typename TConfig::template ClusterArray<TAlpaka>& clusters)
 {
 #if (NDEBUG)
     std::ofstream clusterFile;
     clusterFile.open(path);
     clusterFile << clusters.used << "\n";
-    Cluster* clusterPtr = alpaka::mem::view::getPtrNative(clusters.clusters);
+    typename TConfig::Cluster* clusterPtr =
+        alpaka::mem::view::getPtrNative(clusters.clusters);
 
     DEBUG("writing", clusters.used, "clusters to", path);
 
@@ -27,10 +27,12 @@ void saveClusters(std::string path,
                     << "\n";
 
         // write cluster
-        for (uint8_t y = 0; y < CLUSTER_SIZE; ++y) {
+        for (uint8_t y = 0; y < TConfig::CLUSTER_SIZE; ++y) {
             clusterFile << "\t";
-            for (uint8_t x = 0; x < CLUSTER_SIZE; ++x) {
-                clusterFile << clusterPtr[i].data[x + y * CLUSTER_SIZE] << " ";
+            for (uint8_t x = 0; x < TConfig::CLUSTER_SIZE; ++x) {
+                clusterFile << clusterPtr[i]
+                                   .data[x + y * TConfig::CLUSTER_SIZE]
+                            << " ";
             }
 
             clusterFile << "\n";
@@ -41,15 +43,16 @@ void saveClusters(std::string path,
 #endif
 }
 
-template <typename TBuffer>
+template <typename TConfig, typename TBuffer>
 void save_image(std::string path, TBuffer* data, std::size_t frame_number)
 {
 #if (NDEBUG)
     std::ofstream img;
     img.open(path + ".txt");
-    for (std::size_t j = 0; j < DIMY; j++) {
-        for (std::size_t k = 0; k < DIMX; k++) {
-            double h = double(data[frame_number].data[(j * DIMY) + k]);
+    for (std::size_t j = 0; j < TConfig::DIMY; j++) {
+        for (std::size_t k = 0; k < TConfig::DIMX; k++) {
+            double h = double(
+                data[frame_number].data[(j * TConfig::DIMY) + k]);
             img << h << " ";
         }
[…]

### 2019-04-17 include/jungfrau-photoncounter/Dispenser.hpp

@@ -21,25 +21,40 @@
 #include <iostream>
 #include <limits>
 
-template <typename TAlpaka> class Dispenser {
+template <typename TConfig, template <std::size_t> typename TAccelerator>
+class Dispenser {
 public:
+    // use types defined in the config struct
+    using TAlpaka = TAccelerator<TConfig::MAPSIZE>;
+    using MaskMap = typename TConfig::MaskMap;
+
     /**
      * Dispenser constructor
      * @param Maps-Struct with initial gain
      */
-    Dispenser(FramePackage<GainMap, TAlpaka> gainMap,
-              boost::optional<typename TAlpaka::HostBuf<MaskMap>> mask)
+    Dispenser(typename TConfig::template FramePackage<typename TConfig::GainMap,
+                                                      TAlpaka> gainMap,
+              boost::optional<typename TAlpaka::template HostBuf<MaskMap>> mask)
         : host(alpakaGetHost<TAlpaka>()),
           gain(gainMap),
-          mask((mask ? *mask : alpakaAlloc<MaskMap>(host, SINGLEMAP))),
-          drift(alpakaAlloc<DriftMap>(host, SINGLEMAP)),
-          gainStage(alpakaAlloc<GainStageMap>(host, SINGLEMAP)),
-          maxValueMaps(alpakaAlloc<EnergyMap>(host, DEV_FRAMES)),
+          mask((mask ? *mask
+                     : alpakaAlloc<typename TConfig::MaskMap>(
+                           host,
+                           decltype(TConfig::SINGLEMAP)(TConfig::SINGLEMAP)))),
+          drift(alpakaAlloc<typename TConfig::DriftMap>(
+              host,
+              decltype(TConfig::SINGLEMAP)(TConfig::SINGLEMAP))),
+          gainStage(alpakaAlloc<typename TConfig::GainStageMap>(
+              host,
+              decltype(TConfig::SINGLEMAP)(TConfig::SINGLEMAP))),
+          maxValueMaps(alpakaAlloc<typename TConfig::EnergyMap>(
+              host,
+              decltype(TConfig::DEV_FRAMES)(TConfig::DEV_FRAMES))),
           pedestalFallback(false),
           init(false),
           ringbuffer(TAlpaka::STREAMS_PER_DEV * alpakaGetDevCount<TAlpaka>()),
-          pedestal(PEDEMAPS, host),
-          initPedestal(PEDEMAPS, host),
[…]

### 2019-04-17 include/jungfrau-photoncounter/deviceData.hpp

@@ -11,7 +11,8 @@ enum State { FREE, PROCESSING, READY };
  * devices. It's fully templated to use one of the structs provided
  * by Alpakaconfig.hpp.
  */
-template <typename TAlpaka> struct DeviceData {
+template <typename Config, typename TAlpaka> struct DeviceData {
+
     std::size_t id;
     std::size_t numMaps;
     typename TAlpaka::DevAcc device;
@@ -20,51 +21,85 @@ template <typename TAlpaka> struct DeviceData {
     State state;
 
     // device maps
-    typename TAlpaka::AccBuf<DetectorData> data;
-    typename TAlpaka::AccBuf<GainMap> gain;
-    typename TAlpaka::AccBuf<InitPedestalMap> initialPedestal;
-    typename TAlpaka::AccBuf<PedestalMap> pedestal;
-    typename TAlpaka::AccBuf<MaskMap> mask;
-    typename TAlpaka::AccBuf<DriftMap> drift;
-    typename TAlpaka::AccBuf<GainStageMap> gainStage;
-    typename TAlpaka::AccBuf<GainStageMap> gainStageOutput;
-    typename TAlpaka::AccBuf<EnergyMap> energy;
-    typename TAlpaka::AccBuf<EnergyMap> maxValueMaps;
-    typename TAlpaka::AccBuf<EnergyValue> maxValues;
-    typename TAlpaka::AccBuf<PhotonMap> photon;
-    typename TAlpaka::AccBuf<SumMap> sum;
-    typename TAlpaka::AccBuf<Cluster> cluster;
-    typename TAlpaka::AccBuf<unsigned long long> numClusters;
+    typename TAlpaka::template AccBuf<typename Config::DetectorData> data;
+    typename TAlpaka::template AccBuf<typename Config::GainMap> gain;
+    typename TAlpaka::template AccBuf<typename Config::InitPedestalMap>
+        initialPedestal;
+    typename TAlpaka::template AccBuf<typename Config::PedestalMap> pedestal;
+    typename TAlpaka::template AccBuf<typename Config::MaskMap> mask;
+    typename TAlpaka::template AccBuf<typename Config::DriftMap> drift;
+    typename TAlpaka::template AccBuf<typename Config::GainStageMap> gainStage;
+    typename TAlpaka::template AccBuf<typename Config::GainStageMap>
+        gainStageOutput;
+    typename TAlpaka::template AccBuf<typename Config::EnergyMap> energy;
[…]

### 2019-04-17 include/jungfrau-photoncounter/kernel/Calibration.hpp

@@ -1,7 +1,7 @@
 #pragma once
-#include "../Config.hpp"
 #include "helpers.hpp"
 
+template<typename Config>
 struct CalibrationKernel {
     template <typename TAcc,
               typename TDetectorData,
@@ -16,15 +16,17 @@ struct CalibrationKernel {
                                   TMaskMap* const mask,
                                   TNumFrames const numFrames) const -> void
     {
+        constexpr auto PEDEMAPS = Config::PEDEMAPS;
+        
         auto id = getLinearIdx(acc);
 
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
 
         const std::size_t FRAMESPERSTAGE[] = {
-            FRAMESPERSTAGE_G0, FRAMESPERSTAGE_G1, FRAMESPERSTAGE_G2};
-
+            Config::FRAMESPERSTAGE_G0, Config::FRAMESPERSTAGE_G1, Config::FRAMESPERSTAGE_G2};
+        
         // find expected gain stage
         char expectedGainStage;
         for (int i = 0; i < PEDEMAPS; ++i) {
@@ -50,7 +52,7 @@ struct CalibrationKernel {
                 initPedestal(acc,
                              adc,
                              initPedestalMap[gainStage][id],
-                             MOVING_STAT_WINDOW_SIZE);
+                             Config::MOVING_STAT_WINDOW_SIZE);
             }
             else {
                 // set moving window size for other pedestal stages to the

### 2019-04-17 include/jungfrau-photoncounter/kernel/CheckStdDev.hpp

@@ -1,7 +1,7 @@
 #pragma once
-#include "../Config.hpp"
 #include "helpers.hpp"
 
+template<typename Config>
 struct CheckStdDevKernel {
     template <typename TAcc,
               typename TInitPedestalMap,
@@ -15,12 +15,12 @@ struct CheckStdDevKernel {
         auto id = getLinearIdx(acc);
         
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
 
         // check if measured RMS exceeds threshold
         if (initPedestalMap[0][id].stddev >
-            threshold * threshold * MOVING_STAT_WINDOW_SIZE)
+            threshold * threshold * Config::MOVING_STAT_WINDOW_SIZE)
             mask->data[id] = false;
     }
 };

### 2019-04-17 include/jungfrau-photoncounter/kernel/ClusterFinder.hpp

@@ -1,7 +1,7 @@
 #pragma once
-#include "../Config.hpp"
 #include "helpers.hpp"
 
+template<typename Config>
 struct ClusterFinderKernel {
     template <typename TAcc,
               typename TDetectorData,
@@ -29,15 +29,15 @@ struct ClusterFinderKernel {
                                   TNumFrames const numFrames,
                                   TCurrentFrame const currentFrame,
                                   bool pedestalFallback,
-                                  TNumStdDevs const c = C) const -> void
+                                  TNumStdDevs const c = Config::C) const -> void
     {
         auto id = getLinearIdx(acc);
 
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
 
-        constexpr auto n = CLUSTER_SIZE;
+        constexpr auto n = Config::CLUSTER_SIZE;
 
         if (currentFrame) {
             auto adc = getAdc(detectorData[currentFrame - 1].data[id]);
@@ -46,8 +46,8 @@ struct ClusterFinderKernel {
             decltype(id) max;
             const auto& energy = energyMaps[currentFrame - 1].data[id];
             const auto& stddev = initPedestalMaps[gainStage][id].stddev;
-            if (indexQualifiesAsClusterCenter(id)) {
-                findClusterSumAndMax(
+            if (indexQualifiesAsClusterCenter<Config>(id)) {
+              findClusterSumAndMax<Config>(
                     energyMaps[currentFrame - 1].data, id, sum, max);
 
                 // check cluster conditions
@@ -55,14 +55,14 @@ struct ClusterFinderKernel {
                     id == max) {
                     auto& cluster =
                         getClusterBuffer(acc, clusterArray, numClusters);
-                    copyCluster(energyMaps[currentFrame - 1], id, cluster);
+                    copyCluster<Config>(energyMaps[currentFrame - 1], id, cluster);
                 }
 
                 // check dark pixel condition
                 else if (-c * stddev <= energy && c * stddev >= energy &&
                          !pedestalFallback) {
                     updatePedestal(adc,
-                                   MOVING_STAT_WINDOW_SIZE,
+                                   Config::MOVING_STAT_WINDOW_SIZE,
                                    pedestalMaps[gainStage][id]);
                 }
             }

### 2019-04-17 include/jungfrau-photoncounter/kernel/Conversion.hpp

@@ -1,7 +1,7 @@
 #pragma once
-#include "../Config.hpp"
 #include "helpers.hpp"
 
+template<typename Config>
 struct ConversionKernel {
     template <typename TAcc,
               typename TDetectorData,
@@ -23,12 +23,12 @@ struct ConversionKernel {
                                   TNumFrames const numFrames,
                                   TMask const* const mask,
                                   bool pedestalFallback,
-                                  TNumStdDevs const c = C) const -> void
+                                  TNumStdDevs const c = Config::C) const -> void
     {
         auto id = getLinearIdx(acc);
 
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
 
         for (TNumFrames i = 0; i < numFrames; ++i) {
@@ -54,7 +54,7 @@ struct ConversionKernel {
             if (pedestal - c * stddev <= adc && pedestal + c * stddev >= adc &&
                 !pedestalFallback) {
                 updatePedestal(
-                    adc, MOVING_STAT_WINDOW_SIZE, pedestalMaps[gainStage][id]);
+                    adc, Config::MOVING_STAT_WINDOW_SIZE, pedestalMaps[gainStage][id]);
             }
         }
     }

### 2019-04-17 include/jungfrau-photoncounter/kernel/DriftMap.hpp

@@ -1,7 +1,7 @@
 #pragma once
-#include "../Config.hpp"
 #include "helpers.hpp"
 
+template<typename Config>
 struct DriftMapKernel {
     template <typename TAcc,
               typename TInitPedestalMap,
@@ -16,7 +16,7 @@ struct DriftMapKernel {
         auto id = getLinearIdx(acc);
         
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
 
         driftMaps->data[id] =

### 2019-04-17 include/jungfrau-photoncounter/kernel/GainStageMasking.hpp

@@ -1,7 +1,7 @@
 #pragma once
-#include "../Config.hpp"
 #include "helpers.hpp"
 
+template<typename Config>
 struct GainStageMaskingKernel {
     template <typename TAcc,
               typename TGainStageMap,
@@ -16,13 +16,13 @@ struct GainStageMaskingKernel {
         auto id = getLinearIdx(acc);
         
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
         
         // use masks to check whether the channel is valid or masked out
         bool isValid = !mask ? 1 : mask->data[id];
 
         outputGainStageMaps->data[id] =
-            (isValid ? inputGainStageMaps[numFrame].data[id] : MASKED_VALUE);
+            (isValid ? inputGainStageMaps[numFrame].data[id] : Config::MASKED_VALUE);
     }
 };

### 2019-04-17 include/jungfrau-photoncounter/kernel/GainmapInversion.hpp

@@ -1,8 +1,7 @@
 #pragma once
-
-#include "../Config.hpp"
 #include "helpers.hpp"
 
+template<typename Config>
 struct GainmapInversionKernel {
     template <typename TAcc, typename TGain>
     ALPAKA_FN_ACC auto operator()(TAcc const& acc,
@@ -12,10 +11,10 @@ struct GainmapInversionKernel {
         auto id = getLinearIdx(acc);
         
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
         
-        for (size_t i = 0; i < GAINMAPS; ++i) {
+        for (size_t i = 0; i < Config::GAINMAPS; ++i) {
             gainmaps[i][id] = 1.0 / gainmaps[i][id];
         }
     }

### 2019-04-17 include/jungfrau-photoncounter/kernel/MaxValueCopy.hpp

@@ -1,7 +1,7 @@
 #pragma once
-
 #include <alpaka/alpaka.hpp>
 
+template<typename Config>
 struct MaxValueCopyKernel {
     ALPAKA_NO_HOST_ACC_WARNING
 
@@ -14,7 +14,7 @@ struct MaxValueCopyKernel {
         auto id = getLinearIdx(acc);
         
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
         
         if(id < numFrames) {

### 2019-04-17 include/jungfrau-photoncounter/kernel/PhotonFinder.hpp

@@ -1,7 +1,7 @@
 #pragma once
-#include "../Config.hpp"
 #include "helpers.hpp"
 
+template<typename Config>
 struct PhotonFinderKernel {
     template <typename TAcc,
               typename TDetectorData,
@@ -25,12 +25,12 @@ struct PhotonFinderKernel {
                                   TNumFrames const numFrames,
                                   TMask* const mask,
                                   bool pedestalFallback,
-                                  TNumStdDevs const c = C) const -> void
+                                  TNumStdDevs const c = Config::C) const -> void
     {
         auto id = getLinearIdx(acc);
         
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
 
         for (TNumFrames i = 0; i < numFrames; ++i) {
@@ -60,7 +60,7 @@ struct PhotonFinderKernel {
             auto& photonCount = photonMaps[i].data[id];
 
             // calculate photon count from calibrated energy
-            photonCount = (energy + BEAMCONST) * PHOTONCONST;
+            photonCount = (energy + Config::BEAMCONST) * Config::PHOTONCONST;
 
             const auto& pedestal = pedestalMaps[gainStage][id];
             const auto& stddev = initPedestalMaps[gainStage][id].stddev;
@@ -69,7 +69,7 @@ struct PhotonFinderKernel {
             if (pedestal - c * stddev <= adc && pedestal + c * stddev >= adc &&
                 !pedestalFallback) {
                 updatePedestal(
-                    adc, MOVING_STAT_WINDOW_SIZE, pedestalMaps[gainStage][id]);
+                    adc, Config::MOVING_STAT_WINDOW_SIZE, pedestalMaps[gainStage][id]);
             }
         }
     }

### 2019-04-17 include/jungfrau-photoncounter/kernel/Summation.hpp

@@ -1,5 +1,7 @@
-#include "../Config.hpp"
+#pragma once
+#include "AlpakaHelper.hpp"
 
+template<typename Config>
 struct SummationKernel {
     template <typename TAcc,
               typename TData,
@@ -15,7 +17,7 @@ struct SummationKernel {
         auto id = getLinearIdx(acc);
         
         // check range
-        if (id >= MAPSIZE)
+        if (id >= Config::MAPSIZE)
             return;
         
         for (TNumFrames i = 0; i < numFrames; ++i) {

### 2019-04-17 include/jungfrau-photoncounter/kernel/helpers.hpp

@@ -86,16 +86,19 @@ ALPAKA_FN_ACC ALPAKA_FN_INLINE auto updatePedestal(TAdcValue const adc,
     pedestal += (adc - pedestal) / count;
 }
 
-template <typename TThreadIndex>
+template <typename TConfig, typename TThreadIndex>
 ALPAKA_FN_ACC ALPAKA_FN_INLINE auto
 indexQualifiesAsClusterCenter(TThreadIndex id) -> bool
 {
-    constexpr auto n = CLUSTER_SIZE;
-    return (id % DIMX >= n / 2 && id % DIMX <= DIMX - (n + 1) / 2 &&
-            id / DIMX >= n / 2 && id / DIMX <= DIMY - (n + 1) / 2);
+    // To future self: We checked this multiple times. This should be correct.
+    constexpr auto n = TConfig::CLUSTER_SIZE;
+    return (id % TConfig::DIMX >= n / 2 &&
+            id % TConfig::DIMX <= TConfig::DIMX - (n + 1) / 2 &&
+            id / TConfig::DIMX >= n / 2 &&
+            id / TConfig::DIMX <= TConfig::DIMY - (n + 1) / 2);
 }
 
-template <typename TMap, typename TThreadIndex, typename TSum>
+template <typename TConfig, typename TMap, typename TThreadIndex, typename TSum>
 ALPAKA_FN_ACC ALPAKA_FN_INLINE auto findClusterSumAndMax(TMap const& map,
                                                          TThreadIndex const id,
                                                          TSum& sum,
@@ -104,10 +107,10 @@ ALPAKA_FN_ACC ALPAKA_FN_INLINE auto findClusterSumAndMax(TMap const& map,
 {
     TThreadIndex it = 0;
     max = 0;
-    constexpr auto n = CLUSTER_SIZE;
+    constexpr int n = TConfig::CLUSTER_SIZE;
     for (int y = -n / 2; y < (n + 1) / 2; ++y) {
         for (int x = -n / 2; x < (n + 1) / 2; ++x) {
-            it = id + y * DIMX + x;
+            it = id + y * TConfig::DIMX + x;
             if (map[max] < map[it])
                 max = it;
             sum += map[it];
@@ -126,19 +129,22 @@ getClusterBuffer(TAcc const& acc,
     return clusterArray[i];
 }
 
-template <typename TMap, typename TThreadIndex, typename TCluster>
+template <typename TConfig,
[…]

### 2019-04-17 src/Filecache.cpp

@@ -1,23 +1,0 @@
-#include "Filecache.hpp"
-
-#include <sys/stat.h>
-#include <sys/types.h>
-#include <unistd.h>
-
-
-auto Filecache::getFileSize(const std::string path) const -> off_t
-{
-    struct stat fileStat;
-    stat(path.c_str(), &fileStat);
-    return fileStat.st_size;
-}
-
-Filecache::Filecache(std::size_t size)
-    : buffer([size]() {
-          void* p(malloc(size));
-          return reinterpret_cast<char*>(p);
-             }(), free),
-      bufferPointer(buffer.get()),
-      sizeBytes(size)
-{
-}

### 2019-04-17 src/main.cpp

@@ -10,7 +10,12 @@
  * only change this line to change the backend
  * see Alpakaconfig.hpp for all available
  */
-using Accelerator = GpuCudaRt; // CpuSerial;
+template <std::size_t MAPSIZE>
+using Accelerator = GpuCudaRt<MAPSIZE>;//CpuSerial<MAPSIZE>;//GpuCudaRt<MAPSIZE>;//GpuCudaRt<MAPSIZE>; // CpuSerial;
+using Config = MoenchConfig;
+using ConcreteAcc = Accelerator<Config::MAPSIZE>;
+
+//template<> constexpr std::size_t Config::PEDEMAPS;
 
 auto main(int argc, char* argv[]) -> int
 {
@@ -18,83 +23,88 @@ auto main(int argc, char* argv[]) -> int
     t = Clock::now();
 
     // create a file cache for all input files
-    Filecache* fc = new Filecache(1024UL * 1024 * 1024 * 16);
+    Filecache<Config>* fc = new Filecache<Config>(1024UL * 1024 * 1024 * 16);
     DEBUG("filecache created");
 
     // load maps
-    FramePackage<DetectorData, Accelerator> pedestaldata(
-        fc->loadMaps<DetectorData, Accelerator>(
+    typename Config::FramePackage<typename Config::DetectorData, ConcreteAcc>
+        pedestaldata(fc->loadMaps<typename Config::DetectorData, ConcreteAcc>(
             //"../../data_pool/px_101016/allpede_250us_1243__B_000000.dat",
             "../../../moench_data/"
             "1000_frames_pede_e17050_1_00018_00000.dat",
             true));
     DEBUG(pedestaldata.numFrames, "pedestaldata maps loaded");
 
-    FramePackage<DetectorData, Accelerator> data(
-        fc->loadMaps<DetectorData, Accelerator>(
+    typename Config::FramePackage<typename Config::DetectorData, ConcreteAcc>
+        data(fc->loadMaps<typename Config::DetectorData, ConcreteAcc>(
             //"../../data_pool/px_101016/Insu_6_tr_1_45d_250us__B_000000.dat",
             "../../../moench_data/"
             "e17050_1_00018_00000_image.dat",
             true));
     DEBUG(data.numFrames, "data maps loaded");
 
-    FramePackage<GainMap, Accelerator> gain(
-        fc->loadMaps<GainMap, Accelerator>(
+    typename Config::FramePackage<typename Config::GainMap, ConcreteAcc> gain(
[…]

### 2019-04-17 include/jungfrau-photoncounter/Dispenser.hpp
Cleaned up calcPedestal().

@@ -558,26 +558,11 @@ private:
         dev->state = PROCESSING;
         dev->numMaps = numMaps;
 
-
-        DEBUG("not fsd");
-
-
-
-        //DEBUG("size", dev->data.m_extentElements[0], dev->data.m_extentElements[1], dev->data.m_extentElements[2]);
-
-        std::cout << std::endl;
-
-        
-
         alpakaCopy(
             dev->queue,
             dev->data,
             alpakaViewPlainPtrHost<TAlpaka, TDetectorData>(data, host, numMaps),
             numMaps);
-
-
-        DEBUG("sfdsff");
-
 
         // copy offset data from last initialized device
         auto prevDevice = (nextFull + devices.size() - 1) % devices.size();
@@ -587,36 +572,20 @@ private:
                    devices[prevDevice].pedestal,
                    decltype(TConfig::PEDEMAPS)(TConfig::PEDEMAPS));
 
-
-        DEBUG("sfdsff 2");
-
-
         alpakaCopy(dev->queue,
                    dev->initialPedestal,
                    devices[prevDevice].initialPedestal,
                    decltype(TConfig::PEDEMAPS)(TConfig::PEDEMAPS));
-
-
-        DEBUG("sfdsff 3");
-
 
         alpakaCopy(dev->queue,
                    dev->mask,
                    devices[prevDevice].mask,
                    decltype(TConfig::SINGLEMAP)(TConfig::SINGLEMAP));
 
-
-        DEBUG("sfdsff 4");
-
-
         // increase nextFull and nextFree (because pedestal data isn't
         // downloaded like normal data)
         nextFull = (nextFull + 1) % devices.size();
         nextFree = (nextFree + 1) % devices.size();
-
-
-        DEBUG("sfdsff5");
-
 
         if (!init) {
             alpakaMemSet(dev->queue,
@@ -627,10 +596,6 @@ private:
             init = true;
         }
 
-
-        DEBUG("sfdsff 6");
-
-
         CalibrationKernel<TConfig> calibrationKernel{};
         auto const calibration(
             alpakaCreateKernel<TAlpaka>(getWorkDiv<TAlpaka>(),
@@ -640,10 +605,6 @@ private:
                                         alpakaNativePtr(dev->pedestal),
                                         alpakaNativePtr(dev->mask),
                                         dev->numMaps));
-
-
-        DEBUG("sfdsff7");
-
 
         alpakaEnqueueKernel(dev->queue, calibration);
 

### 2019-04-18 Publication/manuscript.tex
Update manuscript.tex

--- 2019-04-02 Publication/manuscript.tex
@@
 
-\affil[1]{HZDR}
-\affil[2]{TU Dresden}
-\affil[3]{PSI}
+\affil[1]{Helmholtz-Zentrum Dresden -- Rossendorf, Bautzner Landstra{\ss}e 400, 01328 Dresden, Germany}
+\affil[2]{Technische Universit\"at Dresden, 01069 Dresden}
+\affil[3]{Paul Scherrer Institut, 5232 Villigen-PSI,  Switzerland }
 
@@
 
-Increasing data rates during FEL experiments require dedicated detectors as well as advanced methods for fast data processing\\
+The increasing data rates during FEL experiments require dedicated detectors as well as advanced methods for fast data processing.
+At PSI, the Jungfrau detector (\cite{jungfrau1}, \cite{jungfrau2}, \cite{jungfrau3}, \cite{Mozzanica_2016}) was developed which is capable due to a gain switching scheme to register single photons as well as photon bunches in a charge integration mode.
+A conversion of the detector data to the energy and the number of registered photons already during data aquisition is beneficial for the analysis of the measurement and furthermore for the reduction of storage space.
+This online data conversion algorithm has to correct each pixel value using the gain maps and the continously updated pedestal maps following a lab-based calibration  procedure~\cite{jungfrau2}.
+The summation of a couple of frames and the cluster identification in the data maps further reduce the amount of data.
+The different modules of one detector system can be processed fully parallel, however, the data conversion itself is not data parallel.
+Therefore, and due to the need for an advanced parallel implementation which is independent of the computing hardware, an Alpaka~\cite{Matthes17} version was written and its performance was tested.\\
 
-Jungfrau detector (\cite{jungfrau1}, \cite{jungfrau2}, \cite{jungfrau3}): pixel detector with gain switching scheme for large range of photon rates (single pixel to photon bunches)\\
+\textcolor{blue}{TODO: Related work}\\
 
-online data conversion: calculate energy and number of photons from detector response for each pixel using continuously updated pedestal maps\\
-
-Find clusters\\
-
-Hardware-independent computation\\
-
-Different numbers of modules, parallel processing\\
-
-GPU / Alpaka: \cite{Matthes17}: portable, parallel and scalable code\\
-
-Related work\\
-
-...\\
-
-In the following, we describe ....\\
+In section~\ref{subsec:det}, we review the design of Jungfrau detectors, followed by the description of the dedicated data conversion algorithm~\ref{subsec:alg}, its hardware-independent implementation~\ref{subsec:alpaka} and benchmark tests in~\ref{subsec:benchmark}.
+The achieved results and conclusions are presented in sections~\ref{sec:results} and \ref{sec:conclusions}.
 
 \section{Methods}
-\subsection{Abilities and applications of the Jungfrau Detector (PSI)}
-
+\subsection{Abilities and applications of the Jungfrau detector (PSI)}
+\label{subsec:det}
 ...
@@
 \subsection{Data processing algorithm (PSI)}
-conversion from detector data to energy and number of photons\\
+\label{subsec:alg}
+conversion from detector data to energy and number of photons including pixel mask, pedestal calculation, pedestal tracking, pedestal correction, gain correction, conversion to number of photons (in the case of monochromatic incident radiation)\\
 
@@
 \subsection{Alpaka implementation of fast data processing (Jonas, Florian)}
+\label{subsec:alpaka}
 
 \subsection{Benchmark tests (Jonas, Florian)}
+\label{subsec:benchmark}
 Design, objectives, and evaluation of tests
@@
 \section{Results}
+\label{sec:results}
 \subsection{Achieved improvements (Jonas)}
@@
 \section{Conclusions and Outlook}
-
+\label{sec:conclusions}
 Is presented method applicable / useful for other detectors, e.g. AGIPD? \\
@@
 \section{Acknowledges}
-Thanks to Alpaka developers.....\\
+The authors would like to thank  the Alpaka developers (\textcolor{blue}{names?})for their support.\\
 

### 2019-04-26 Publication/manuscript.tex
First draft of the JF sections from PSI.
Suggest mentioning Alpaka in the title.
First draft of the abstract, sections 1 and 2 from Sophie, Martin, Carlos and Aldo.
Introduction will need merging with HZDR paragraph.
Few references added, and a few renamed for clarity.
We are not sure how long the final article should be. This work could be too detailed, in which case we can cut down.

@@
 \begin{document}
-\title{Fast processing of Jungfrau detector data}
+\title{Fast processing of JUNGFRAU data with the Alpaka library}
 \author{Jonas Schenke$^{1,2}$, 
@@
 Michael Bussmann$^1$,
-Carlos Lopez-Cuenza$^3$,
+Carlos Lopez-Cuenca$^3$,
 Aldo Mozzanica$^3$,\\
@@
 \affil[2]{Technische Universit\"at Dresden, 01069 Dresden}
-\affil[3]{Paul Scherrer Institut, 5232 Villigen-PSI,  Switzerland }
+\affil[3]{Paul Scherrer Institut, 5232 Villigen-PSI,  Switzerland}
 
@@
 {\bf Abstract}\\
-...text...\\
+Highly segmented photon detectors play a crucial role at synchrotron and X-ray free electron laser facilities, detecting the photons scattered from targets of interest.
+These measurements allow the determination of biological and inorganic structures and processes, leading to an advanced understanding of life and material sciences and enabling improvements and discoveries in pharmaceutical and engineering industries.
+However, as targets become ever-more challenging and photon fluxes intensify, detector frame rates increase and the generated data streams risk becoming overwhelming.
+Here we demonstrate one solution, exploiting the highly parallelisable nature of data from the JUNGFRAU charge integrating photon detector and processing using the Alpaka library.
+We find that...
+To date, no other solution was able to process JUNGFRAU data at this throughput(?).
+Our results demonstrate that parallel processing, for example in a GPU, is a valuable tool in the treatment of photon detector data.
+Advantages of this method include online data correction aiding storage efficiency by allowing raw data to be immediately discarded and enabling fast feedback to users for whom time and target resources are at a premium.
+We see this as an important step in the development of a chain of specialised hardware and software tools necessary to operate charge integrating photon detectors at high frame rates.
 
@@
 
+Data rates and volumes at photon science facilities are increasing due to many technological advancements, including higher beam intensities, robotic or serial sample delivery systems, and photon detectors capable of ever increasing frame rates.
+Charge integrating detectors compound the issue, as the raw detector output must be processed before the number of photons detected is retrieved.
+This requires advanced methods for fast, online data processing without which data volumes threaten becoming a limiting factor, quickly filling storage capacity, overwhelming computing processing power and slowing feedback to users.
+
+The JUNGFRAU detector is one such charge integrating photon detector~\cite{Mozzanica_2016}, which has been designed for SwissFEL, the free electron laser at the Paul Scherrer Institut in Switzerland~\cite{SwissFEL}.
+Two 16 Mpixel JUNGFRAU detectors were installed at SwissFEL in 2018 and are now operational~\cite{JFapplications}.
+Due to the repetition rate and pulsed illumination of SwissFEL, the detectors are operated continuously with short ($10\;\mu\mathrm{s}$) integration time and at a frame rate of 100 Hz, meaning that each 16 Mpixel detector generates a reasonable data rate of 3.2 GB/s.
+
+However, JUNGFRAU has also proved advantageous under challenging conditions at synchrotrons~\cite{Leonarski_2018} where the illumination is continuous, requiring a full duty cycle from the detector and hence much higher frame rates.
+Recent experimental setups have included a one Mpixel detector operated continuously at 1.1 kHz (2.3 GB/s) with online visualisation but slow offline data treatment~\cite{Leonarski_2018} and a four Mpixel detector operated at 1.1 kHz (9.2 GB/s) for short periods of time ($<200$ s) without visualisation but with fast data processing straight after acquisition (Japan, no ref yet).
+Our goal is to operate a 16 Mpixel detector continuously at 2.2 kHz, successfully receiving the raw data at 74 GB/s, providing an online visualisation, processing it online to convert the raw data into number of photons per pixel, and saving to disk.
+This necessitates specialised capability for receiving, processing and file writing, motivating this work.
+
 The increasing data rates during FEL experiments require dedicated detectors as well as advanced methods for fast data processing.
-At PSI, the Jungfrau detector (\cite{jungfrau1}, \cite{jungfrau2}, \cite{jungfrau3}, \cite{Mozzanica_2016}) was developed which is capable due to a gain switching scheme to register single photons as well as photon bunches in a charge integration mode.
+At PSI, the Jungfrau detector (\cite{JFapplications}, \cite{JFcalibration}, \cite{JFoperation}, \cite{Mozzanica_2016}) was developed which is capable due to a gain switching scheme to register single photons as well as photon bunches in a charge integration mode.
 A conversion of the detector data to the energy and the number of registered photons already during data aquisition is beneficial for the analysis of the measurement and furthermore for the reduction of storage space.
-This online data conversion algorithm has to correct each pixel value using the gain maps and the continously updated pedestal maps following a lab-based calibration  procedure~\cite{jungfrau2}.
+This online data conversion algorithm has to correct each pixel value using the gain maps and the continously updated pedestal maps following a lab-based calibration  procedure~\cite{JFcalibration}.
 The summation of a couple of frames and the cluster identification in the data maps further reduce the amount of data.
@@
 \section{Methods}
-\subsection{Abilities and applications of the Jungfrau detector (PSI)}
+\subsection{JUNGFRAU detector description (PSI)}
 \label{subsec:det}
-...
+JUNGFRAU is a hybrid photon detector consisting of a $320\;\mu\mathrm{m}$ thick silicon sensor in which photons are absorbed and converted into an electrical signal, and an Application Specific Integrated Circuit (ASIC) chip in which the signal generated by the photons in the silicon is processed.
+Both sensor and ASIC layers are pixelated with $75\;\mu\mathrm{m}$ pitch, to provide two dimensional information of the position of the incident photons.
+
+The JUNGFRAU ASIC is realised with UMC 110 nm CMOS technology and operates in the `charge integrating' mode, meaning that the signal received by the ASIC within a well-defined, configurable time window (an exposure) is integrated and amplified~\cite{Mozzanica_2016}.
+Charge integrating detectors typically suffer from either a low dynamic range or poor resolution in the single photon regime.
+JUNGFRAU achieves both high dynamic range and single photon resolution by employing a gain switching technique, whereby the amplification factor of the initial photon-induced signal is dependent on the size of the signal itself.
+The gain switching happens automatically and independently per pixel and per exposure.
+The gain factor employed by each pixel (high, medium or low) in each exposure is encoded in a 2-bit output.
+
+Each sensor has the dimensions of $1030\times514$ pixels, resulting in an active area of $7.7\times3.8$ cm$^2$.
+Eight $256 \times 256$ pixel ASICs are bump-bonded to each sensor in $4\times2$ arrangement to achieve an electrical connection in every pixel.
+Between the ASICs are two rows or columns of double-sized pixels, ensuring no gaps or loss of sensitivity.
+The eight ASICs are glued to a high density interconnect (HDI) and electrically connected with wire-bonds along the two long sides.
+Additional wire-bonds to the sensor provide the high voltage to bias the sensor, crucial for detecting the charge generated there by the absorbed photons.
+A readout board connects to the back of the HDI to collect the signals processed by the eight ASICs.
+Four 14-bit ADCs convert the incoming analogue signal to digital samples.
+An FPGA arranges this 14-bit signal information from each pixel together with the 2-bit gain information of each pixel into a 1 MB image (a frame) of each exposure.
+This is output as UDP packets through two 10 Gb optical fibres.
+
+This single detector module can either operate alone, or can be tiled together with other detector modules to form part of a larger detector system, with typical sizes ranging from 1 Mpixel (two modules) to 16 Mpixels (32 modules).
+Figure~\ref{fig:jfdetector} shows the front and back of a 4 Mpixel (8 module) detector.
+Each module retains its individual control (Ethernet), data (10 Gb optical fibres) and power connections.
+The only inter-module connection is a flat-ribbon cable used for triggering purposes.
+Data rates thus scale linearly with the number of modules.
+
+While sending the UDP packets in parallel with 2$\times$10~Gb/s per module is less challenging, receiving them in one or a few spots for further processing is more complex.
+Implementing this on a standard hardware with a non-optimised Linux system easily results in packet loss due to CPU and memory bandwidth limitations.
+Several avenues to resolve this issue are under development.
+Offloading the CPU can be achieved on a software level by bypassing the network stack, which avoids several memory copies.
+Another option is to offload the full receiving process into an FPGA, which receives the UDP packets with it's 100~Gb/s interface, reassembles full images and sends them over PCIe directly into host via DMA or GPU memory over GPUDirect.
+This goes beyond the scope of the current work, which focuses on processing the successfully received JUNGFRAU data.
 
@@
 \includegraphics[width=0.30\textwidth]{jungfraudetector.jpg}
-\caption{Jungfrau detector}
+\hfill
+\includegraphics[width=0.50\textwidth]{JF4M_back.jpg}
+\caption{A 4 Mpixel JUNGFRAU detector shown from the front (left) and back (right).
+Photons impinge on the eight mirrored sensor surfaces, which are tiled to ensure minimal dead area.
+The readout boards plug into the back of the modules.
+Each has its own Ethernet (purple cables) and optical fibre (green cables).
+Water cooling (blue pipes) and power (black cable) are provided in common.
+A flat ribbon cable, visible on the left, connects to all readout boards for triggering purposes.}
 \label{fig:jfdetector}
@@
 
+\subsection{Raw data treatment (PSI)}
+
+To retrieve the information of how much energy, and in monochromatic situations, how many photons were detected in each pixel in each frame, a series of corrections must be applied to the raw 16-bit output of each pixel.
+
+\subsubsection{Pedestal correction and tracking}
+As any charge integrating detector, the JUNGFRAU detects some current in every pixel during every exposure, even if no photons were absorbed.
+This dark current or `pedestal' is individual to each pixel, and is strongly dependent on the configuration of the detector, for instance the exposure time, and on environmental factors such as temperature.
+It is therefore measured by acquiring a series of dark frames before each measurement with photons.
+To measure the pedestals in medium and low gain, the detector is forced to switch gain during the dark frames.
+The ADC values recorded in the dark frames in high, medium and low gain are averaged to give a per-pixel, per-gain pedestal.
+A rolling average is preferred [ref cook], both to avoid committing long arrays of data to memory, and to place higher importance on more recent exposures.
+
+In the case of steady operation at XFELs, with a short integration time and the detector at room temperature, recording the pedestals every few hours is sufficient.
+However in the case of operation at synchrotrons, with a long integration time and the detector chilled to $\sim{-10}^{\circ}C$, the pedestals are much more sensitive~\cite{JFoperation}.
+They must be recorded before every measurement with photons, and moreover, must be tracked through the measurement.
+This is achieved by filtering the recorded exposures, and updating the pedestal measurement of a particular pixel if the frame recorded a value within $\pm2\sigma$ of the existing pedestal.
+This avoids exposures containing photon signals (typically $>10\sigma$ from the noise) from biasing the pedestal.
+
+\subsubsection{Gain correction}
+The pedestal-corrected ADC samples are translated into energies, so taking into account the amplification of the original signal by the ASIC.
+The amplification factors are calculated per pixel and per gain in a dedicated laboratory based procedure during the commissioning of the detector~\cite{JFcalibration}.
+Once the energy detected per pixel has been retrieved, in monochromatic situations this can be converted to a number of photons by normalising by the energy of the incident radiation, as shown in the following equation:
+\begin{equation}
+    N_{\gamma} = \frac{\mathrm{ADC\;output\;[ADU]} - \mathrm{pedestal\;[ADU]}}{\mathrm{gain\;[ADU/keV]} \times E_{\gamma} \mathrm{[keV]}}
+\end{equation}
+where $N_{\gamma}$ is the number of photons detected, $E_{\gamma}$ is the photon energy, and the pedestal and gain corrections are pixel dependent.
+
+\subsubsection{Pixel mask}
+A pixel mask can be defined using information from the calibration procedure and the pedestal calculation.
+Pixels can be non-responsive, or respond badly, for a number of reasons, including a chip level defect, bump bonding defect, wire-bond column defect or high leakage current.
+A pixel which could not be calibrated for any of these reasons, or could not have an accurate pedestal calculated, is masked in the later processing.
+
+\subsubsection{Frame summation}
+Depending on the application, further processing may be necessary.
+For macro-molecular crystallography, groups of subsequent frames (in our experience from 2 to 100 frames) can be summed together before saving~\cite{Leonarski_2018}.
+Depending on the rotation speed of the sample, this can improve later analysis to determine crystal structure.
+It also acts as an effective compression of the data.
+
+\subsubsection{Pixel clustering}
+In the case of soft X-rays at low incoming flux, pixels may be clustered to draw additional positioning information from the data. Description needed~\cite{Cartier_2015}.
+
 \subsection{Data processing algorithm (PSI)}
 \label{subsec:alg}
-conversion from detector data to energy and number of photons including pixel mask, pedestal calculation, pedestal tracking, pedestal correction, gain correction, conversion to number of photons (in the case of monochromatic incident radiation)\\
-
- summation of frames\\
-
- clustering (reference?)\\
+The following enumerates the typical data acquisition and processing steps for a single measurement with photons in a synchrotron environment, which is the most challenging in terms of data rate and which requires pedestal tracking.
+\begin{enumerate}
+\item acquisition of 1000 dark frames in forced switching to medium gain. The medium gain pedestal is calculated and pixels not in medium gain are added to the pixel mask.
+\item acquisition of 1000 dark frames in forced switching to low gain. The low gain pedestal is calculated and pixels not in low gain are added to the pixel mask.
+\item acquisition of at least 1000 dark frames in gain switching mode.
+Using the first 1000 frames, the high gain pedestal is calculated and pixels not in high gain are added to the pixel mask.
+At 1000 frames, pixels with larger than normal noise (pedestal RMS) are also added to the pixel mask and the noise level per pixel in high gain is taken as the pedestal tracking threshold.
+After 1000 frames, the high gain pedestal is tracked and the detector and software is `ready and waiting for photons'.
+\item without stopping and restarting the detector, the beamline shutter opens and frames with photons are acquired in gain switching mode.
+Frames must be converted to number of photons per pixel.
+High gain pedestals are tracked.
+Frames can be optionally summed or clustered before being saved.
+\end{enumerate}
+Whether online or offline, whether implemented in CPU, GPU, FPGA or other hardware, this processing pipeline must be realised in order to convert the raw data output from the JUNGFRAU into information in terms of number of photons per pixel, for use in later analyses.
 

### 2019-07-03 include/bench.hpp
Implemented first test benchmark using google benchmark. Also fixes #57.

@@ -1,0 +1,296 @@
+#pragma once
+
+#include "jungfrau-photoncounter/Alpakaconfig.hpp"
+#include "jungfrau-photoncounter/Config.hpp"
+#include "jungfrau-photoncounter/Dispenser.hpp"
+
+#include "Filecache.hpp"
+#include "jungfrau-photoncounter/Debug.hpp"
+
+// a struct to hold the input parameters for the benchmarking function
+template <typename Config, typename TAccelerator> struct BenchmarkingInput {
+    using MaskMap = typename Config::MaskMap;
+
+    // input data
+    typename Config::template FramePackage<typename Config::DetectorData,
+                                           TAccelerator>
+        pedestalData;
+    typename Config::template FramePackage<typename Config::DetectorData,
+                                           TAccelerator>
+        data;
+    typename Config::template FramePackage<typename Config::GainMap,
+                                           TAccelerator>
+        gain;
+    boost::optional<typename TAccelerator::template HostBuf<MaskMap>> maskPtr;
+
+    // output buffers
+    boost::optional<
+        typename Config::template FramePackage<typename Config::EnergyMap,
+                                               TAccelerator>>
+        energy;
+    boost::optional<
+        typename Config::template FramePackage<typename Config::PhotonMap,
+                                               TAccelerator>>
+        photons;
+    boost::optional<
+        typename Config::template FramePackage<typename Config::SumMap,
+                                               TAccelerator>>
+        sum;
+    typename Config::template ClusterArray<TAccelerator>* clusters;
+    boost::optional<
+        typename Config::template FramePackage<typename Config::EnergyValue,
+                                               TAccelerator>>
+        maxValues;
+    typename Config::ExecutionFlags ef;
+
+    // constructor
+    BenchmarkingInput(
+        typename Config::template FramePackage<typename Config::DetectorData,
+                                               TAccelerator> pedestalData,
+        typename Config::template FramePackage<typename Config::DetectorData,
+                                               TAccelerator> data,
+        typename Config::template FramePackage<typename Config::GainMap,
+                                               TAccelerator> gain,
+        boost::optional<typename TAccelerator::template HostBuf<MaskMap>>
+            maskPtr,
+        boost::optional<
+            typename Config::template FramePackage<typename Config::EnergyMap,
+                                                   TAccelerator>> energy,
[…]

### 2019-07-03 include/jungfrau-photoncounter/Config.hpp

@@ -187,7 +187,7 @@ typedef std::chrono::high_resolution_clock Clock;
 typedef std::chrono::milliseconds ms;
 static Clock::time_point t;
 
-#ifdef NDEBUG
+#ifdef VERBOSE
 #include <iostream>
 
 // empty print to end recursion

### 2019-07-03 src/bench.cpp

@@ -1,313 +1,21 @@
-//#include <benchmark/benchmark.h>
+#include <benchmark/benchmark.h>
+
+#include "bench.hpp"
 
 #include "jungfrau-photoncounter/Alpakaconfig.hpp"
 #include "jungfrau-photoncounter/Config.hpp"
 #include "jungfrau-photoncounter/Dispenser.hpp"
 
-#include "Filecache.hpp"
 #include "jungfrau-photoncounter/Debug.hpp"
 
-
-#include <type_traits>
-
-
-// a struct to hold the input parameters for the benchmarking function
-template <typename Config, typename TAccelerator> struct BenchmarkingInput {
-    using MaskMap = typename Config::MaskMap;
-
-    // input data
-    typename Config::template FramePackage<typename Config::DetectorData,
-                                           TAccelerator>
-        pedestalData;
-    typename Config::template FramePackage<typename Config::DetectorData,
-                                           TAccelerator>
-        data;
-    typename Config::template FramePackage<typename Config::GainMap,
-                                           TAccelerator>
-        gain;
-    boost::optional<typename TAccelerator::template HostBuf<MaskMap>> maskPtr;
-
-    // output buffers
-    boost::optional<
-        typename Config::template FramePackage<typename Config::EnergyMap,
-                                               TAccelerator>>
-        energy;
-    boost::optional<
-        typename Config::template FramePackage<typename Config::PhotonMap,
-                                               TAccelerator>>
-        photons;
-    boost::optional<
-        typename Config::template FramePackage<typename Config::SumMap,
-                                               TAccelerator>>
-        sum;
-    typename Config::template ClusterArray<TAccelerator>*
-        clusters;
-    boost::optional<
-        typename Config::template FramePackage<typename Config::EnergyValue,
-                                               TAccelerator>>
-        maxValues;
-    typename Config::ExecutionFlags ef;
-
-    // constructor
-    BenchmarkingInput(
-        typename Config::template FramePackage<typename Config::DetectorData,
-                                               TAccelerator> pedestalData,
-        typename Config::template FramePackage<typename Config::DetectorData,
-                                               TAccelerator> data,
-        typename Config::template FramePackage<typename Config::GainMap,
-                                               TAccelerator> gain,
-        boost::optional<typename TAccelerator::template HostBuf<MaskMap>>
-            maskPtr,
[…]

### 2019-07-03 src/main.cpp

@@ -10,8 +10,7 @@
  * only change this line to change the backend
  * see Alpakaconfig.hpp for all available
  */
-template <std::size_t MAPSIZE>
-using Accelerator = GpuCudaRt<MAPSIZE>;
+template <std::size_t MAPSIZE> using Accelerator = GpuCudaRt<MAPSIZE>;
 using Config = JungfrauConfig;
 using ConcreteAcc = Accelerator<Config::MAPSIZE>;
 
@@ -121,19 +120,21 @@ auto main(int argc, char* argv[]) -> int
 
     // process data maps
     while (downloaded < data.numFrames) {
-      // save the upload future
+        // save the upload future
         uploadFutures.emplace_back(dispenser.uploadData(data, offset, ef));
 
-        // get the number of frames that have been uploaded and add them to the offset
+        // get the number of frames that have been uploaded and add them to the
+        // offset
         offset = std::get<0>(*uploadFutures.rbegin());
 
         // save the download future
         downloadFutures.emplace_back(
             dispenser.downloadData(energy, photon, sum, maxValues, clusters));
 
-        // update the number of downloaded frames, if any frames where downloaded 
-        if (currently_downloaded_frames =
-                std::get<0>(*downloadFutures.rbegin())) {
+        // update the number of downloaded frames, if any frames where
+        // downloaded
+        currently_downloaded_frames = std::get<0>(*downloadFutures.rbegin());
+        if (currently_downloaded_frames) {
             downloaded += currently_downloaded_frames;
             DEBUG(downloaded,
                   "/",

### 2019-07-03 Publication/manuscript.tex
Added some related works.

@@
 \usepackage{lipsum}% <- For dummy text
-\definechangesauthor[name={Heide}, color=blue]{Heide}
-\setremarkmarkup{(#2)}
+%\definechangesauthor[name={Heide}, color=blue]{Heide}
+%\definechangesauthor[name={Jonas}, color=pink]{Jonas}
+%\setremarkmarkup{(#2)}
 
@@
 \affil[1]{Helmholtz-Zentrum Dresden -- Rossendorf, Bautzner Landstra{\ss}e 400, 01328 Dresden, Germany}
-\affil[2]{Technische Universit\"at Dresden, 01069 Dresden}
+\affil[2]{Technische Universit\"at Dresden, 01069 Dresden, Germany}
 \affil[3]{Paul Scherrer Institut, 5232 Villigen-PSI,  Switzerland}
@@
 At PSI, the Jungfrau detector (\cite{JFapplications}, \cite{JFcalibration}, \cite{JFoperation}, \cite{Mozzanica_2016}) was developed which is capable due to a gain switching scheme to register single photons as well as photon bunches in a charge integration mode.
-A conversion of the detector data to the energy and the number of registered photons already during data aquisition is beneficial for the analysis of the measurement and furthermore for the reduction of storage space.
-This online data conversion algorithm has to correct each pixel value using the gain maps and the continously updated pedestal maps following a lab-based calibration  procedure~\cite{JFcalibration}.
+A conversion of the detector data to the energy and the number of registered photons already during data acquisition is beneficial for the analysis of the measurement and furthermore for the reduction of storage space.
+This online data conversion algorithm has to correct each pixel value using the gain maps and the continuously updated pedestal maps following a lab-based calibration  procedure~\cite{JFcalibration}.
 The summation of a couple of frames and the cluster identification in the data maps further reduce the amount of data.
 The different modules of one detector system can be processed fully parallel, however, the data conversion itself is not data parallel.
-Therefore, and due to the need for an advanced parallel implementation which is independent of the computing hardware, an Alpaka~\cite{Matthes17} version was written and its performance was tested.\\
+Therefore, and due to the need for an advanced parallel implementation which is independent of the computing hardware, an Alpaka \cite{Matthes17} version was written and its performance was tested.\\
 
 \textcolor{blue}{TODO: Related work}\\
+\textcolor{pink}{
+common preprocessing steps:
+ * offset correction \& flat field correction [DSSC \cite{moch2014calibration}, JF \cite{JFcalibration}, EIGER\cite{tinti2018electron} / \cite{dinapoli2013eiger}, MOENCH \cite{cartier2016micrometer}, [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]]
+ * gain correction [DSSC \cite{moch2014calibration}, EIGER \cite{tinti2018electron} / \cite{dinapoli2013eiger}, MOENCH \cite{cartier2016micrometer}]
+ * dynamic gain switching [AGIPD \cite{allahgholi2015agipd}, GOTTHARD \cite{mozzanica2012gotthard}, JF \cite{JFcalibration}, [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
+ * pixel sorting (on FPGA)[AGPID \cite{becker2013high}]
+ * discard uninteresting events (in readeout circuit)[DSSC \cite{erdinger2012dssc}]
+ * photon counting [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
+ * charge summing / interpolation (to counter charge sharing; on ASIC)[MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
+ * cluster finder [MOENCH \cite{cartier2016micrometer}]
+ * sub pixel interpolation [MOENCH \cite{cartier2016micrometer}]}\\
+%TODO: common calibration and preprocessing steps / online preprocessing -> FPGAs
+%TODO: offline preprocessing -> XDS, ...
+%TODO: further analysis -> crystallography
+%TODO: related acceleration libraries
 
@@
 
+requirements:
+ - realtime data processing (at least 1 detector at 2.5 kHz)
+ - simple and extensible design
+ - portable to other architectures (CPU OpenMP, TBB, ...) with minimal code changes
+
+optional requirements:
+ - easily adaptable to similar detectors (e.g. MÖNCH)
+
+design overview:
+ - wrap for most important alpaka functions (for ease of use)
+ - templated for maximal performance and configurability
+ - stream processing of data: upload data to GPU -> process -> download again
+ - multiple detectors and multiple GPUs usable in parallel
+
 \subsection{Benchmark tests (Jonas, Florian)}
@@
 Design, objectives, and evaluation of tests
+
+benchmarked with google benchmark
 

### 2020-01-06 compile_cuda.sh
Minor bugfixes, benchmarks should now be runnable

@@ -1,0 +1,20 @@
+#! /bin/bash
+#SBATCH --job-name=CUDA_Compile
+#SBATCH --partition=gpu
+#SBATCH --time=23:00:00
+#SBATCH --nodes=1
+#SBATCH --ntasks=4
+#SBATCH --mail-type=ALL
+#SBATCH --mail-user=j.schenke@hzdr.de
+#SBATCH -o log.slurm-%j.out
+#SBATCH -e err.slurm-%j.out
+
+set -x
+
+export MODULEPATH=$MODULEPATH:/home/schenk24/tools/hemera_spack/spack/share/spack/modules/linux-centos7-x86_64/
+module load git gcc cmake cuda boost/1.68.0 alpaka-develop-gcc-6.4.0-iu4hm3r 
+
+mkdir build_cuda
+cd build_cuda
+cmake .. -DCMAKE_BUILD_TYPE=Release -DALPAKA_CUDA_ARCH=60 -DBENCHMARKING_ENABLED=ON
+make -j

### 2020-01-06 include/bench.hpp

@@ -84,13 +84,13 @@ auto setUp(ExecutionFlags flags, std::string pedestalPath, std::string gainPath,
         fc.template loadMaps<typename Config::MaskMap, ConcreteAcc>(maskPath);
     DEBUG(mask.numFrames, "mask maps loaded");
   }
-
+  
   // create empty, optional input mask
   using MaskMap = typename Config::MaskMap;
   tl::optional<typename ConcreteAcc::template HostBuf<MaskMap>> maskPtr;
   if (mask.numFrames == Config::SINGLEMAP)
     maskPtr = mask.data;
-
+  
   // allocate space for output data
   FramePackage<typename Config::EnergyMap, ConcreteAcc> energy_data(
       data.numFrames);
@@ -125,7 +125,7 @@ auto setUp(ExecutionFlags flags, std::string pedestalPath, std::string gainPath,
     sum = sum_data;
   if (flags.maxValue)
     maxValues = maxValues_data;
-
+  
   // return configuration
   return BenchmarkingInput<Config, ConcreteAcc>(
       pedestalData, data, gain, beamConst, maskPtr, energy, photon, sum,
@@ -135,7 +135,7 @@ auto setUp(ExecutionFlags flags, std::string pedestalPath, std::string gainPath,
 // calibrate the detector
 template <typename Config, template <std::size_t> typename Accelerator>
 auto calibrate(const BenchmarkingInput<Config, Accelerator<Config::MAPSIZE>>
-                   &benchmarkingConfig) -> Dispenser<Config, Accelerator> {
+                   &benchmarkingConfig) -> Dispenser<Config, Accelerator> {  
   Dispenser<Config, Accelerator> dispenser(benchmarkingConfig.gain,
                                            benchmarkingConfig.beamConst,
                                            benchmarkingConfig.maskPtr);
@@ -170,20 +170,6 @@ auto bench(
 
   typename Config::template ClusterArray<ConcreteAcc> *clusters =
       benchmarkingConfig.clusters;
-
-  save_image<Config>("pedestal_begin0",
-                     alpakaNativePtr(dispenser.downloadPedestaldata().data), 0);
-
-  // download and save std dev of the initial pedestal map
-  auto initPed = dispenser.downloadInitialPedestaldata();
-  std::ofstream outPede("pede_stddev.bin", std::ios::binary);
-  InitPedestal *pedePtr = alpakaNativePtr(initPed.data)->data;
[…]

### 2020-01-06 include/jungfrau-photoncounter/Config.hpp

@@ -310,8 +310,6 @@ typedef std::chrono::high_resolution_clock Clock;
 typedef std::chrono::milliseconds ms;
 static Clock::time_point t;
 
-#define VERBOSE
-
 #ifdef VERBOSE
 #include <iostream>
 

### 2020-01-06 include/jungfrau-photoncounter/Dispenser.hpp

@@ -322,7 +322,7 @@ public:
     // try uploading one data package
     if (offset + TConfig::DEV_FRAMES <= data.numFrames) {
 
-      // TODO: combine these two cases
+      //! @todo: combine these two cases
       auto dataView = data.getView(offset, TConfig::DEV_FRAMES);
 
       auto result = processData(dataView, TConfig::DEV_FRAMES, flags, energy,

### 2020-01-06 run.py

@@ -1,0 +1,103 @@
+#!/bin/python
+
+import os
+import sys
+from itertools import product
+
+# define possible values
+masking = [0, 1]
+devFrames = [10, 100, 1000]
+maxValue = [0, 1]
+summation = [0, 2, 10, 20, 100]
+clusterSizes = [2, 3, 7, 11]
+clusterFiles = ["JF", "CL0", "CL4", "CL8"]
+energyFiles = ["JF", "G0", "G13"]
+
+# create photon and energy configurations
+configurations = list(product(summation, devFrames, [0], [0, 1], masking, maxValue, energyFiles))
+
+# add clustering configurations
+configurations += list(product([0], devFrames, clusterSizes, [2], masking, [0], clusterFiles))
+
+# remove impossible configurations
+configurations = [(s, d, c, m, mask, mv, files) for (s, d, c, m, mask, mv, files) in configurations if s < d]
+
+# read configuration index
+configID = int(sys.argv[1])
+
+# get config
+(sumFrames, devFramesOption, clusterSize, mode, maskingEnable, maxValueEnable, inputFile) = configurations[configID]
+
+# initialize values
+summationEnable = 0 if sumFrames == 0 else 1
+benchmarkID = sumFrames * len(devFrames) * len(clusterSizes) + devFramesOption * len(clusterSizes) + clusterSize
+    
+iterationCount = 100
+beamConst = 12.4
+
+# extract file pathes
+pedestalPath = ""
+gainPath = ""
+dataPath = ""
+
+if inputFile == "JF": # jungfrau data
+    pedestalPath = "../../../data_pool/px_101016/allpede_250us_1243__B_000000.dat"
+    gainPath = "../../../data_pool/px_101016/gainMaps_M022.bin"
+    dataPath = "../../../data_pool/px_101016/Insu_6_tr_1_45d_250us__B_000000.dat"
+elif inputFile == "G0": # g0 data
+    pedestalPath = "../../../data_pool/synthetic/pede.bin"
+    gainPath = "../../../data_pool/px_101016/gainMaps_M022.bin"
+    dataPath = "../../../data_pool/synthetic/g0.bin"
+elif inputFile == "G13": # g13 data
+    pedestalPath = "../../../data_pool/synthetic/pede.bin"
+    gainPath = "../../../data_pool/px_101016/gainMaps_M022.bin"
[…]

### 2020-01-06 run_cuda.sh

@@ -1,0 +1,26 @@
+#!/bin/bash
+#SBATCH --job-name=CUDA_Run_1
+#SBATCH --partition=gpu
+#SBATCH --time=1:00:00
+#SBATCH --nodes=1
+#SBATCH --ntasks=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --gres=gpu:2
+#SBATCH --mem=300000
+#SBATCH --cpus-per-task=24
+#SBATCH --exclusive
+#SBATCH --array=0-360
+#SBATCH --mail-type=ALL
+#SBATCH --mail-user=j.schenke@hzdr.de
+#SBATCH -o log.slurm-cuda_%A_%a.out
+#SBATCH -e err.slurm-cuda_%A_%a.out
+
+set -x
+
+export CUDA_VISIBLE_DEVICES=0
+
+export MODULEPATH=$MODULEPATH:/home/schenk24/tools/hemera_spack/spack/share/spack/modules/linux-centos7-x86_64/
+module load git gcc cmake cuda boost/1.68.0 alpaka-develop-gcc-6.4.0-iu4hm3r python 
+
+cd build_cuda
+python ../run.py $SLURM_ARRAY_TASK_ID

### 2020-01-06 scripts/test_data_generator.cpp

@@ -262,18 +262,11 @@ void generateCluster(std::string path, float clusterAmount,
 }
 
 int main() {
-  // generateCalibration("/bigdata/hplsim/production/jungfrau-photoncounter/tmp/pede.bin");
-  // generateMainG0("/bigdata/hplsim/production/jungfrau-photoncounter/tmp/g0.bin",
-  // 10); //working
-  // generateMainG13("/bigdata/hplsim/production/jungfrau-photoncounter/tmp/g13.bin",
-  // 10); //working
-  // generateCluster<3>(
-  //    "/bigdata/hplsim/production/jungfrau-photoncounter/tmp/cluster_0.bin",
-  //    0.f, 10); //working
-  // generateCluster<3>("/bigdata/hplsim/production/jungfrau-photoncounter/tmp/cluster_4.bin",
-  // 0.04f, 10); // seems working
-  generateCluster<3>(
-      "/bigdata/hplsim/production/jungfrau-photoncounter/tmp/cluster_8.bin",
-      0.08f, 10); // seems working
+  generateCalibration("/bigdata/hplsim/production/jungfrau-photoncounter/data_pool/synthetic/pede.bin");
+  generateMainG0("/bigdata/hplsim/production/jungfrau-photoncounter/data_pool/synthetic/g0.bin", 10000);
+  generateMainG13("/bigdata/hplsim/production/jungfrau-photoncounter/data_pool/synthetic/g13.bin", 10000);
+  generateCluster<3>("/bigdata/hplsim/production/jungfrau-photoncounter/data_pool/synthetic/cluster_0.bin", 0.f, 10000);
+  generateCluster<3>("/bigdata/hplsim/production/jungfrau-photoncounter/data_pool/synthetic/cluster_4.bin", 0.04f, 10000);
+  generateCluster<3>("/bigdata/hplsim/production/jungfrau-photoncounter/data_pool/synthetic/cluster_8.bin", 0.08f, 10000);
   return 0;
 }

### 2020-01-06 src/bench.cpp

@@ -17,22 +17,19 @@
  */
 
 template <std::size_t TMapSize>
-using Accelerator =
-    GpuCudaRt<TMapSize>; // CpuOmp2Blocks<TMapSize>; // CpuSerial<TMapSize>;
-                         // // GpuCudaRt<TMapSize>;
+using Accelerator = GpuCudaRt<TMapSize>; // CpuOmp2Blocks<TMapSize>;
+// CpuSerial<TMapSize>;
+// GpuCudaRt<TMapSize>;
 
 constexpr auto framesPerStageG0 = Values<std::size_t, 1000>();
 constexpr auto framesPerStageG1 = Values<std::size_t, 1000>();
 constexpr auto framesPerStageG2 = Values<std::size_t, 999>();
 constexpr auto dimX = Values<std::size_t, 1024>();
 constexpr auto dimY = Values<std::size_t, 512>();
-constexpr auto sumFrames =
-    Values<std::size_t, 2>(); // Values<std::size_t, 2, 10, 20, 100>();
-constexpr auto devFrames =
-    Values<std::size_t, 10>(); // Values<std::size_t, 10, 100, 1000>();
+constexpr auto sumFrames = Values<std::size_t, 2, 10, 20, 100>();
+constexpr auto devFrames = Values<std::size_t, 10, 100, 1000>();
 constexpr auto movingStatWindowSize = Values<std::size_t, 100>();
-constexpr auto clusterSize =
-    Values<std::size_t, 2, 3>(); // Values<std::size_t, 2, 3, 7, 11>();
+constexpr auto clusterSize = Values<std::size_t, 2, 3, 7, 11>();
 constexpr auto cs = Values<std::size_t, 5>();
 
 constexpr auto parameterSpace =
@@ -40,167 +37,188 @@ constexpr auto parameterSpace =
     sumFrames * devFrames * movingStatWindowSize * clusterSize * cs;
 
 template <class Tuple> struct ConfigFrom {
-  using G0 = Get_t<0, Tuple>;
-  using G1 = Get_t<1, Tuple>;
-  using G2 = Get_t<2, Tuple>;
-  using DimX = Get_t<3, Tuple>;
-  using DimY = Get_t<4, Tuple>;
-  using SumFrames = Get_t<5, Tuple>;
-  using DevFrames = Get_t<6, Tuple>;
-  using MovingStatWinSize = Get_t<7, Tuple>;
-  using ClusterSize = Get_t<8, Tuple>;
[…]

### 2020-01-06 Publication/manuscript.tex
Added new bullet points to implementation section of the paper.

--- 2019-04-26 Publication/manuscript.tex
@@
 \usepackage{lipsum}% <- For dummy text
-\definechangesauthor[name={Heide}, color=blue]{Heide}
-\setremarkmarkup{(#2)}
+%\definechangesauthor[name={Heide}, color=blue]{Heide}
+%\definechangesauthor[name={Jonas}, color=magenta]{Jonas}
+%\setremarkmarkup{(#2)}
 
@@
 
-{\bf Keywords:} Photon pixel detector, fast data processing, GPU programming, Alpaka...
+{\bf Keywords:} Photon pixel detector, fast data processing, GPU programming, Alpaka \textcolor{magenta}{...}
 
@@
 At PSI, the Jungfrau detector (\cite{JFapplications}, \cite{JFcalibration}, \cite{JFoperation}, \cite{Mozzanica_2016}) was developed which is capable due to a gain switching scheme to register single photons as well as photon bunches in a charge integration mode.
-A conversion of the detector data to the energy and the number of registered photons already during data aquisition is beneficial for the analysis of the measurement and furthermore for the reduction of storage space.
-This online data conversion algorithm has to correct each pixel value using the gain maps and the continously updated pedestal maps following a lab-based calibration  procedure~\cite{JFcalibration}.
+A conversion of the detector data to the energy and the number of registered photons already during data acquisition is beneficial for the analysis of the measurement and furthermore for the reduction of storage space.
+This online data conversion algorithm has to correct each pixel value using the gain maps and the continuously updated pedestal maps following a lab-based calibration  procedure~\cite{JFcalibration}.
 The summation of a couple of frames and the cluster identification in the data maps further reduce the amount of data.
 The different modules of one detector system can be processed fully parallel, however, the data conversion itself is not data parallel.
-Therefore, and due to the need for an advanced parallel implementation which is independent of the computing hardware, an Alpaka~\cite{Matthes17} version was written and its performance was tested.\\
+Therefore, and due to the need for an advanced parallel implementation which is independent of the computing hardware, an Alpaka \cite{Matthes17} version was written and its performance was tested.\\
 
 \textcolor{blue}{TODO: Related work}\\
+\textcolor{magenta}{
+common pre-processing steps:
+\begin{itemize}
+	\item offset correction \& flat field correction [DSSC \cite{moch2014calibration}, JF \cite{JFcalibration}, EIGER\cite{tinti2018electron} / \cite{dinapoli2013eiger}, MOENCH \cite{cartier2016micrometer}, [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]]
+	\item gain correction [DSSC \cite{moch2014calibration}, EIGER \cite{tinti2018electron} / \cite{dinapoli2013eiger}, MOENCH \cite{cartier2016micrometer}]
+	\item dynamic gain switching [AGIPD \cite{allahgholi2015agipd}, GOTTHARD \cite{mozzanica2012gotthard}, JF \cite{JFcalibration}, [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
+	\item pixel sorting (on FPGA)[AGPID \cite{becker2013high}]
+	\item discard uninteresting events (in readeout circuit)[DSSC \cite{erdinger2012dssc}]
+	\item photon counting [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
+	\item charge summing / interpolation (to counter charge sharing; on ASIC)[MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
+	\item cluster finder [MOENCH \cite{cartier2016micrometer}]
+	\item sub pixel interpolation [MOENCH \cite{cartier2016micrometer}]
+\end{itemize}
+detector data processing libraries (with pre-processing stage):
+\begin{itemize}
+	\item Karabo \cite{fangohr2018data}
+	\begin{itemize}
+		\item developed to support experiments at the European XFEL
+		\item written in C++ and Python
+		\item pipelined design
+		\item supports offline and online operation and an additional rapid-feedback mode
+		\item can utilize multiple nodes
+		\item also implements file IO and a GUI
+		\item extensible through Karabo-Bridge
+		\item has OnDA interface
+		\item supports Jupytier notebook
+	\end{itemize}
+	\item Hummingbird \cite{daurer2016hummingbird}
+	\begin{itemize}
+		\item optimized for real-time usage
+		\item implemented in Python
+		\item main purpose: analysis of diffraction data at LCLS
+		\item achieves rates above 100 Hz
+		\item contains a GUI and supports multiple file streams at once
+		\item can utilize multiple computing nodes
+		\item can read data generated by psana
+	\end{itemize}
+	\item OnDA \cite{mariani2016onda}
+	\begin{itemize}
+		\item supports online \& offline operation
+		\item can utilize multiple computing nodes (via ZeroMQ and MPI)
+		\item master/worker architecture
+		\item one program instance per core
+		\item designed to give fast online feedback
+		\item main objective: real-time monitoring of X-ray imaging experiments
+		\item easily adaptable
+		\item contains a GUI
+		\item implemented in Python
+		\item can utilize a psana back end
+	\end{itemize}
+	\item psana \cite{damiani2016linac}
+	\begin{itemize}
+		\item = Photon Science ANAlysis
+		\item similar functionality to pyana (which is written entirely in Python), but implemented mostly in C++ with a Python interface
+		\item used to analyze data produced by SLAC National Laboratory
+		\item can utilize multiple compute nodes (via MPI)
+		\item supports online \& offline operation
+		\item contains detector calibration algorithms (e.g. pedestal calibration, bad-pixel determination, etc.) and GUI front end
+		\item designed to give fast online feedback
+	\end{itemize}
+	\item Cheetah \cite{barty2014cheetah}
+	\begin{itemize}
+		\item mode for XFEL applications
+		\item offline \& online processing
+		\item evaluates data quality metrics (e.g. number of Bragg peaks etc.)
+		\item retain only frames with high likelihood of being usable
+		\item --> main task: data reduction
+		\item additional pre-processing
+		\item also has a GUI
+		\item implemented in C++
+	\end{itemize}
+	\item CASS \cite{foucar2012cass}
+	\begin{itemize}
+		\item evaluate data acquired using the CFEL ASG Multi Purpose instrument (CAMP), specifically data collected at FLASH and the European XFEL
+		\item supports multiple detectors in online and offline scenarios
+		\item implemented in C++ with a Qt front end
+		\item consists of input ringbuffer and processing modules (which do pre- \& post-processing)
+		\item GUI implemented in two additional programs: joCASSView (plots data) \& luCASSView (interactive)
+		\item GUIs communicate with computation nodes via TCP/IP
+	\end{itemize}
+	\item XDS \cite{kabsch2010xds}
+	\begin{itemize}
+		\item in development since 1988
+		\item mainly deals with crystallography
+		\item also does pre-processing steps
+		\item no real-time/on-line capability
+	\end{itemize}
+\end{itemize}
+initialized data further used e.g. in crystallography \cite{kabsch2010integration} \cite{white2012crystfel} \cite{grosse2002computational}}\\
 
@@
 
+\textcolor{magenta}{
+\textbf{requirements:}
+\begin{itemize}
+	\item real-time data processing (at least 1 detector at 2.5 kHz)
+	\item simple and extensible design(?)
+	\item portable to other architectures (CPU OpenMP, TBB, ...) with minimal code changes
+\end{itemize}
+optional requirements:
+\begin{itemize}
+	\item easily adaptable to similar detectors (e.g. MÖNCH)
+\end{itemize}
+expected bottlenecks:
+\begin{itemize}
+	\item communication speed between host and GPU
+	\item RAM speed
+\end{itemize}
+\textbf{design overview:}
+\begin{itemize}
+	\item wrap for most important alpaka functions (for ease of use) --$>$ explain alpaka shortly (?)
+	\item templated for maximal performance and configurability (?)
+	\item written in modern C++ for optimal performance
+	\item pipelined stream processing of data: upload data to GPU -$>$ process -$>$ download again
+	\item multiple detectors and multiple GPUs usable in parallel
+\end{itemize}
+$<$insert data flow diagram here$>$ \\
+architecture specific optimizations:
+\begin{itemize}
+	\item contiguous memory vs. grid striding loop
+	\item only upload data if needed
+\end{itemize}
+\textbf{software architecture:}
+\textit{hardware accelerator concept}
+\begin{itemize}
+	\item host
+	\item device
+	\item memory not necessarily shared
+	\item data might need to be copied onto device
+	\item device starts (possibly asynchronous) execution
+	\item result is copied back to host
+	\item common host: CPU
+	\item common devices: GPUs, FPGAs, Multicore CPUs (in general multi- and manycore systems)
+\end{itemize}
+$<$insert hardware sketch here$>$
+\textit{the dispenser}
+\begin{itemize}
+	\item handles devices (initialization, memory management, etc.)
+	\item enqueues data to next free devices (asynchronously)
+	\item copies data back from devices (asynchronously)
+\end{itemize}
+$<$insert dispenser structure here$>$
+\textit{the kernel structure}
+\begin{itemize}
+	\item written in empty struct as static function
+	\item all needed values passed to function, no real access to member variables
+	\item most data passed by pointer, only single numbers passed directly
+	\item kernel has access to device properties (block size, grid size, device specific functions, etc.)
+\end{itemize}
+$<$insert example here$>$
+\textit{easy hardware abstraction}
+\begin{itemize}
+	\item useful settings defined for every possible device class
+	\item device is selected by using statement (similar to a typedef)
+	\item no additional changes needed
+\end{itemize}
+$<$insert example here$>$
+}
+
 \subsection{Benchmark tests (Jonas, Florian)}

### 2020-02-24 include/jungfrau-photoncounter/Alpakaconfig.hpp
Added reference implementation.

@@ -179,7 +179,7 @@ template <std::size_t MAPSIZE> struct CpuSerial {
     template <typename T>
     using HostView = alpaka::mem::view::ViewSubView<DevHost, T, Dim, Size>;
 
-    static constexpr std::size_t STREAMS_PER_DEV = 4;
+    static constexpr std::size_t STREAMS_PER_DEV = 1u;
     static constexpr Size elementsPerThread = 1u;
     static constexpr Size threadsPerBlock = 1u;
     static constexpr Size blocksPerGrid = MAPSIZE;

### 2020-02-24 include/jungfrau-photoncounter/Config.hpp

@@ -9,177 +9,158 @@
 #include "CheapArray.hpp"
 
 // a struct to hold all detector specific configuration variables
-template <std::size_t TFramesPerStageG0,
-          std::size_t TFramesPerStageG1,
-          std::size_t TFramesPerStageG2,
-          std::size_t TDimX,
-          std::size_t TDimY,
-          std::size_t TSumFrames,
-          std::size_t TDevFrames,
-          std::size_t TMovingStatWindowSize,
-          std::size_t TClusterSize,
+template <std::size_t TFramesPerStageG0, std::size_t TFramesPerStageG1,
+          std::size_t TFramesPerStageG2, std::size_t TDimX, std::size_t TDimY,
+          std::size_t TSumFrames, std::size_t TDevFrames,
+          std::size_t TMovingStatWindowSize, std::size_t TClusterSize,
           std::size_t TC>
 struct DetectorConfig {
-    static constexpr std::size_t FRAMESPERSTAGE_G0 = TFramesPerStageG0;
-    static constexpr std::size_t FRAMESPERSTAGE_G1 = TFramesPerStageG1;
-    static constexpr std::size_t FRAMESPERSTAGE_G2 = TFramesPerStageG2;
-    static constexpr std::size_t DIMX = TDimX;
-    static constexpr std::size_t DIMY = TDimY;
-    static constexpr std::size_t SUM_FRAMES = TSumFrames;
-    static constexpr std::size_t DEV_FRAMES = TDevFrames;
-    static constexpr std::size_t MOVING_STAT_WINDOW_SIZE =
-        TMovingStatWindowSize;
-    static constexpr std::size_t CLUSTER_SIZE = TClusterSize;
-    static constexpr std::size_t C = TC;
+  static constexpr std::size_t FRAMESPERSTAGE_G0 = TFramesPerStageG0;
+  static constexpr std::size_t FRAMESPERSTAGE_G1 = TFramesPerStageG1;
+  static constexpr std::size_t FRAMESPERSTAGE_G2 = TFramesPerStageG2;
+  static constexpr std::size_t DIMX = TDimX;
+  static constexpr std::size_t DIMY = TDimY;
+  static constexpr std::size_t SUM_FRAMES = TSumFrames;
+  static constexpr std::size_t DEV_FRAMES = TDevFrames;
+  static constexpr std::size_t MOVING_STAT_WINDOW_SIZE = TMovingStatWindowSize;
[…]

### 2020-02-24 include/jungfrau-photoncounter/Dispenser.hpp

@@ -122,6 +122,51 @@ public:
         while (offset <= data.numFrames - TConfig::DEV_FRAMES) {
             offset += calcPedestaldata(alpakaNativePtr(data.data) + offset,
                                        TConfig::DEV_FRAMES);
+
+
+
+            if(offset >= 990 && offset < 1000) {
+              // download and save std dev of the initial pedestal map
+              auto initPed = downloadInitialPedestaldata();
+              std::ofstream outPede("pede/stddev_" + std::to_string(offset) + ".txt");
+              typename TConfig::InitPedestal *pedePtr = alpakaNativePtr(initPed.data)->data;
+              for (unsigned int y = 0; y < TConfig::DIMY; ++y) {
+                outPede << "\t";
+                for (unsigned int x = 0; x < TConfig::DIMX; ++x) {
+                  outPede << pedePtr[y * TConfig::DIMX + x].stddev << " ";
+                }
+                outPede << "\n";
+              }
+              outPede.flush();
+              outPede.close();
+
+              outPede.open("pede/m_" + std::to_string(offset) + ".txt");
+              pedePtr = alpakaNativePtr(initPed.data)->data;
+              for (unsigned int y = 0; y < TConfig::DIMY; ++y) {
+                outPede << "\t";
+                for (unsigned int x = 0; x < TConfig::DIMX; ++x) {
+                  outPede << pedePtr[y * TConfig::DIMX + x].m << " ";
+                }
+                outPede << "\n";
+              }
+              outPede.flush();
+              outPede.close();
+
+              outPede.open("pede/m2_" + std::to_string(offset) + ".txt");
+              pedePtr = alpakaNativePtr(initPed.data)->data;
+              for (unsigned int y = 0; y < TConfig::DIMY; ++y) {
+                outPede << "\t";
+                for (unsigned int x = 0; x < TConfig::DIMX; ++x) {
+                  outPede << pedePtr[y * TConfig::DIMX + x].m2 << " ";
+                }
+                outPede << "\n";
+              }
+              outPede.flush();
+              outPede.close();
+            }
+
+
+            
         }
 
         // upload remaining frames

### 2020-02-24 include/jungfrau-photoncounter/kernel/Calibration.hpp

@@ -1,8 +1,7 @@
 #pragma once
 #include "helpers.hpp"
 
-template<typename Config>
-struct CalibrationKernel {
+template <typename Config> struct CalibrationKernel {
     template <typename TAcc,
               typename TDetectorData,
               typename TInitPedestalMap,
@@ -17,16 +16,17 @@ struct CalibrationKernel {
                                   TNumFrames const numFrames) const -> void
     {
         constexpr auto PEDEMAPS = Config::PEDEMAPS;
-        
+
         auto id = getLinearIdx(acc);
 
         // check range
         if (id >= Config::MAPSIZE)
             return;
 
-        const std::size_t FRAMESPERSTAGE[] = {
-            Config::FRAMESPERSTAGE_G0, Config::FRAMESPERSTAGE_G1, Config::FRAMESPERSTAGE_G2};
-        
+        const std::size_t FRAMESPERSTAGE[] = {Config::FRAMESPERSTAGE_G0,
+                                              Config::FRAMESPERSTAGE_G1,
+                                              Config::FRAMESPERSTAGE_G2};
+
         // find expected gain stage
         char expectedGainStage;
         for (int i = 0; i < PEDEMAPS; ++i) {

### 2020-02-24 include/jungfrau-photoncounter/kernel/ClusterFinder.hpp

@@ -1,83 +1,67 @@
 #pragma once
 #include "helpers.hpp"
 
-template<typename Config>
-struct ClusterFinderKernel {
-    template <typename TAcc,
-              typename TDetectorData,
-              typename TGainMap,
-              typename TInitPedestalMap,
-              typename TPedestalMap,
-              typename TGainStageMap,
-              typename TEnergyMap,
-              typename TClusterArray,
-              typename TNumClusters,
-              typename TMask,
-              typename TNumFrames,
-              typename TCurrentFrame,
-              typename TNumStdDevs = int>
-    ALPAKA_FN_ACC auto operator()(TAcc const& acc,
-                                  TDetectorData const* const detectorData,
-                                  TGainMap const* const gainMaps,
-                                  TInitPedestalMap* const initPedestalMaps,
-                                  TPedestalMap* const pedestalMaps,
-                                  TGainStageMap* const gainStageMaps,
-                                  TEnergyMap* const energyMaps,
-                                  TClusterArray* const clusterArray,
-                                  TNumClusters* const numClusters,
-                                  TMask* const mask,
-                                  TNumFrames const numFrames,
-                                  TCurrentFrame const currentFrame,
-                                  bool pedestalFallback,
-                                  TNumStdDevs const c = Config::C) const -> void
-    {
-        auto id = getLinearIdx(acc);
+template <typename Config> struct ClusterFinderKernel {
+  template <typename TAcc, typename TDetectorData, typename TGainMap,
+            typename TInitPedestalMap, typename TPedestalMap,
+            typename TGainStageMap, typename TEnergyMap, typename TClusterArray,
+            typename TNumClusters, typename TMask, typename TNumFrames,
+            typename TCurrentFrame, typename TNumStdDevs = int>
+  ALPAKA_FN_ACC auto operator()(
+      TAcc const &acc, TDetectorData const *const detectorData,
+      TGainMap const *const gainMaps, TInitPedestalMap *const initPedestalMaps,
+      TPedestalMap *const pedestalMaps, TGainStageMap *const gainStageMaps,
+      TEnergyMap *const energyMaps, TClusterArray *const clusterArray,
+      TNumClusters *const numClusters, TMask *const mask,
+      TNumFrames const numFrames, TCurrentFrame const currentFrame,
[…]

### 2020-02-24 include/jungfrau-photoncounter/kernel/helpers.hpp

@@ -3,201 +3,175 @@
 #include <alpaka/alpaka.hpp>
 
 template <typename TAcc>
-ALPAKA_FN_ACC ALPAKA_FN_INLINE auto getLinearIdx(const TAcc& acc)
-    -> std::uint64_t
-{
-    auto const globalThreadIdx = alpakaGetGlobalThreadIdx(acc);
-    auto const globalThreadExtent = alpakaGetGlobalThreadExtent(acc);
+ALPAKA_FN_ACC ALPAKA_FN_INLINE auto getLinearIdx(const TAcc &acc)
+    -> std::uint64_t {
+  auto const globalThreadIdx = alpakaGetGlobalThreadIdx(acc);
+  auto const globalThreadExtent = alpakaGetGlobalThreadExtent(acc);
 
-    auto const linearizedGlobalThreadIdx =
-        alpakaGetGlobalLinearizedGlobalThreadIdx(globalThreadIdx,
-                                                 globalThreadExtent);
+  auto const linearizedGlobalThreadIdx =
+      alpakaGetGlobalLinearizedGlobalThreadIdx(globalThreadIdx,
+                                               globalThreadExtent);
 
-    return linearizedGlobalThreadIdx[0u];
+  return linearizedGlobalThreadIdx[0u];
 }
 
 template <typename TDataword>
-ALPAKA_FN_ACC ALPAKA_FN_INLINE auto getAdc(TDataword dataword) -> uint16_t
-{
-    return dataword & 0x3fff;
+ALPAKA_FN_ACC ALPAKA_FN_INLINE auto getAdc(TDataword dataword) -> uint16_t {
+  return dataword & 0x3fff;
 }
 
 template <typename TDataword>
-ALPAKA_FN_ACC ALPAKA_FN_INLINE auto getGainStage(TDataword dataword) -> uint8_t
-{
-    auto g = (dataword & 0xc000) >> 14;
-    // map gain stages 0, 1, 3 to 0, 1, 2
-    if (g == 3)
-        g = 2;
-    return g;
+ALPAKA_FN_ACC ALPAKA_FN_INLINE auto getGainStage(TDataword dataword)
+    -> uint8_t {
+  auto g = (dataword & 0xc000) >> 14;
+  // map gain stages 0, 1, 3 to 0, 1, 2
+  if (g == 3)
+    g = 2;
+  return g;
 }
 
 template <typename TInputMap, typename TOutputMap>
[…]

### 2020-02-24 src/main.cpp

@@ -6,162 +6,424 @@
 
 #include "jungfrau-photoncounter/Debug.hpp"
 
+#include <fstream>
+
 /**
  * only change this line to change the backend
  * see Alpakaconfig.hpp for all available
  */
-template <std::size_t MAPSIZE> using Accelerator = GpuCudaRt<MAPSIZE>;
+template <std::size_t MAPSIZE>
+using Accelerator =
+    GpuCudaRt<MAPSIZE>; // CpuOmp2Blocks<MAPSIZE>; // GpuCudaRt<MAPSIZE>;
 using Config = JungfrauConfig;
 using ConcreteAcc = Accelerator<Config::MAPSIZE>;
 
-auto main(int argc, char* argv[]) -> int
-{
-    // t is used in all debug-messages
-    t = Clock::now();
+auto main(int argc, char *argv[]) -> int {
+  // t is used in all debug-messages
+  t = Clock::now();
 
-    // create a file cache for all input files
-    Filecache<Config>* fc = new Filecache<Config>(1024UL * 1024 * 1024 * 16);
-    DEBUG("filecache created");
+  // create a file cache for all input files
+  Filecache<Config> *fc = new Filecache<Config>(1024UL * 1024 * 1024 * 16);
+  DEBUG("filecache created");
 
-    // load maps
-    typename Config::FramePackage<typename Config::DetectorData, ConcreteAcc>
-        pedestaldata(fc->loadMaps<typename Config::DetectorData, ConcreteAcc>(
-            "../../../data_pool/px_101016/allpede_250us_1243__B_000000.dat",
-            //"../../../moench_data/"
-            //"1000_frames_pede_e17050_1_00018_00000.dat",
-            true));
-    DEBUG(pedestaldata.numFrames, "pedestaldata maps loaded");
+  // load maps
+  typename Config::FramePackage<typename Config::DetectorData, ConcreteAcc>
+      pedestaldata(fc->loadMaps<typename Config::DetectorData, ConcreteAcc>(
+          "../../../data_pool/px_101016/allpede_250us_1243__B_000000.dat",
+          //"../../../moench_data/"
+          //"1000_frames_pede_e17050_1_00018_00000.dat",
+          true));
+  DEBUG(pedestaldata.numFrames, "pedestaldata maps loaded");
 
-    typename Config::FramePackage<typename Config::DetectorData, ConcreteAcc>
[…]

### 2020-02-24 include/bench.hpp
Fixed minor bugs. Still not working correctly.

@@ -171,6 +171,21 @@ auto bench(
   typename Config::template ClusterArray<ConcreteAcc> *clusters =
       benchmarkingConfig.clusters;
 
+
+
+
+
+  //! @todo: remove debugging code
+  // write out initial pedestal data
+  save_image<Config>(
+                     "pedestal_start",
+                     alpakaNativePtr(dispenser.downloadPedestaldata().data), 0);
+
+
+
+
+
+  
   // process data maps
   while (offset < benchmarkingConfig.data.numFrames) {
 
@@ -201,6 +216,32 @@ auto bench(
     futures.emplace_back(dispenser.process(benchmarkingConfig.data, offset,
                                            benchmarkingConfig.ef, energy,
                                            photons, sum, maxValues, clusters));
+
+
+
+
+
+
+
+
+    //! @todo: remove debugging code
+    // write out pedestal and image values
+            std::get<1>(futures.back()).wait();
+        dispenser.synchronize();
+
+        save_image<Config>("pedestal_" + std::to_string(offset),
+            alpakaNativePtr(dispenser.downloadPedestaldata().data), 0);
+
+        if (energy)
+          save_image<Config>("energy_" + std::to_string(offset),
+                             alpakaNativePtr(energy->data), 0);
+
+
+
+
+
+
+        
     auto offset_diff = std::get<0>(*futures.rbegin()) - offset;
     offset = std::get<0>(*futures.rbegin());
     sum_offset += (offset_diff + Config::SUM_FRAMES - 1) / Config::SUM_FRAMES;

### 2020-02-24 include/check.hpp

@@ -61,7 +61,11 @@ bool checkResult(tl::optional<T> result, std::string referencePath) {
   if (readBytes < dataSize || !reference_file.read(&dummyByte, 1).eof()) {
     std::cerr << "Number of read bytes does not match (" << readBytes << " vs. "
               << dataSize << "; eof: " << reference_file.eof() << ")!\n";
-    return false;
+
+
+
+    //! @todo: remove debugging code
+    //return false;
   }
 
   // compare data

### 2020-02-24 include/jungfrau-photoncounter/Alpakaconfig.hpp

@@ -84,7 +84,7 @@ template <std::size_t MAPSIZE> struct CpuTbbBlocks {
   template <typename T>
   using HostView = alpaka::mem::view::ViewSubView<DevHost, T, Dim, Size>;
 
-  static constexpr std::size_t STREAMS_PER_DEV = 4;
+  static constexpr std::size_t STREAMS_PER_DEV = 1u;
   static constexpr Size elementsPerThread = 1u;
   static constexpr Size threadsPerBlock = 1u;
   static constexpr Size blocksPerGrid = MAPSIZE;
@@ -141,9 +141,9 @@ template <std::size_t MAPSIZE> struct CpuOmp2Blocks {
   using HostView = alpaka::mem::view::ViewSubView<DevHost, T, Dim, Size>;
 
   static constexpr std::size_t STREAMS_PER_DEV = 1;
-  static constexpr Size elementsPerThread = 1u;//256u;
+  static constexpr Size elementsPerThread = 256u;
   static constexpr Size threadsPerBlock = 1u;
-  static constexpr Size blocksPerGrid = MAPSIZE;//(MAPSIZE + 255) / 256;
+  static constexpr Size blocksPerGrid = (MAPSIZE + 255) / 256;
 };
 
 template <typename T, typename TBuf, typename... TArgs>
@@ -178,7 +178,7 @@ template <std::size_t MAPSIZE> struct CpuOmp4 {
   template <typename T>
   using HostView = alpaka::mem::view::ViewSubView<DevHost, T, Dim, Size>;
 
-  static constexpr std::size_t STREAMS_PER_DEV = 4;
+  static constexpr std::size_t STREAMS_PER_DEV = 1u;
   static constexpr Size elementsPerThread = 1u;
   static constexpr Size threadsPerBlock = 1u;
   static constexpr Size blocksPerGrid = MAPSIZE;
@@ -216,9 +216,9 @@ template <std::size_t MAPSIZE> struct CpuSerial {
   using HostView = alpaka::mem::view::ViewSubView<DevHost, T, Dim, Size>;
 
   static constexpr std::size_t STREAMS_PER_DEV = 1;
-  static constexpr Size elementsPerThread = 1u;//256u;
+  static constexpr Size elementsPerThread = 256u;
   static constexpr Size threadsPerBlock = 1u;
-  static constexpr Size blocksPerGrid = MAPSIZE;//(MAPSIZE + 255) / 256;
[…]

### 2020-02-24 include/jungfrau-photoncounter/Config.hpp

@@ -410,5 +410,5 @@ void debugPrint(const char* file, unsigned int line, TArgs... args)
 
 // predefine detector configurations
 using JungfrauConfig =
-    DetectorConfig<1000, 1000, 999, 1024, 512, 2, 1, 100, 3, 5>;
+    DetectorConfig<1000, 1000, 999, 1024, 512, 2, 1, 100, 2, 5>;
 using MoenchConfig = DetectorConfig<1000, 0, 0, 400, 400, 10, 300, 100, 3, 5>;

### 2020-02-24 include/jungfrau-photoncounter/Dispenser.hpp

@@ -60,11 +60,11 @@ public:
           deviceContainer(alpakaGetDevs<TAlpaka>())
     {
         initDevices();
-        
+
         // make room for live mask information
         if (!mask) {
-            alpakaMemSet(devices[devices.size() - 1].queue,
-                         devices[devices.size() - 1].mask,
+            alpakaMemSet(devices[0].queue,
+                         devices[0].mask,
                          1,
                          decltype(TConfig::SINGLEMAP)(TConfig::SINGLEMAP));
         }
@@ -117,8 +117,8 @@ public:
         initDevices();
 
         // clear mask
-        alpakaMemSet(devices[devices.size() - 1].queue,
-                     devices[devices.size() - 1].mask,
+        alpakaMemSet(devices[0].queue,
+                     devices[0].mask,
                      1,
                      decltype(TConfig::SINGLEMAP)(TConfig::SINGLEMAP));
 
@@ -127,7 +127,7 @@ public:
     }
 
     /**
-     * Tries to upload all data packages requiered for the inital offset.
+     * Tries to upload all data packages requiered for the initl offset.
      * Only stops after all data packages are uploaded.
      * @param Maps-Struct with datamaps
      * @param stdDevThreshold An standard deviation threshold above which pixels
@@ -144,6 +144,53 @@ public:
         while (offset <= data.numFrames - TConfig::DEV_FRAMES) {
             offset += calcPedestaldata(alpakaNativePtr(data.data) + offset,
                                        TConfig::DEV_FRAMES);
+
+            
+            /*
+            //! @todo: remove debugging code
+            if(offset >= 990 && offset < 1000) {
+                
+              // download and save std dev of the initial pedestal map                                                                                                                                                                                                   
+              auto initPed = downloadInitialPedestaldata();
+              std::ofstream outPede("pede/stddev_" + std::to_string(offset) + ".txt");
+              InitPedestal *pedePtr = alpakaNativePtr(initPed.data)->data;
+              for (unsigned int y = 0; y < TConfig::DIMY; ++y) {
+                outPede << "\t";
+                for (unsigned int x = 0; x < TConfig::DIMX; ++x) {
+                  outPede << pedePtr[y * TConfig::DIMX + x].stddev << " ";
+                }
+                outPede << "\n";
+              }
+              outPede.flush();
+              outPede.close();
[…]

### 2020-02-24 include/jungfrau-photoncounter/deviceData.hpp

@@ -92,7 +92,7 @@ template <typename Config, typename TAlpaka> struct DeviceData {
           cluster(alpakaAlloc<typename Config::Cluster>(
               *device,
               decltype(Config::MAX_CLUSTER_NUM_USER)(
-                  Config::MAX_CLUSTER_NUM_USER) * 10 *
+                  Config::MAX_CLUSTER_NUM_USER) *
                   decltype(Config::DEV_FRAMES)(Config::DEV_FRAMES))),
           numClusters(alpakaAlloc<unsigned long long>(
               *device,

### 2020-02-24 include/jungfrau-photoncounter/kernel/Calibration.hpp

@@ -32,8 +32,8 @@ template <typename Config> struct CalibrationKernel {
           expectedGainStage = i;
           break;
         }
-      }
-
+      }      
+      
       // determine expected gain stage
       for (TNumFrames i = 0; i < numFrames; ++i) {
         if (initPedestalMap[expectedGainStage][id].count ==

### 2020-02-24 include/jungfrau-photoncounter/kernel/ClusterFinder.hpp

@@ -17,19 +17,18 @@ template <typename Config, typename AccConfig> struct ClusterFinderKernel {
       bool pedestalFallback, TNumStdDevs const c = Config::C) const -> void {
     unsigned long globalId = getLinearIdx(acc);
     unsigned long elementsPerThread = getLinearElementExtent(acc);
-    //AccConfig::elementsPerThread; // getLinearElementExtent(acc);
+      //AccConfig::elementsPerThread; // getLinearElementExtent(acc);
 
     constexpr auto n = Config::CLUSTER_SIZE;
 
     // iterate over all elements in the thread
-    //for (auto id = globalId * elementsPerThread;
-    //     id < (globalId + 1) * elementsPerThread; ++id) {
-    auto id = globalId;
+    for (auto id = globalId * elementsPerThread;
+         id < (globalId + 1) * elementsPerThread; ++id) {
 
       // check range
       if (id >= Config::MAPSIZE)
         return;
-
+      
       if (currentFrame) {
         auto adc = getAdc(detectorData[currentFrame - 1].data[id]);
         const auto &gainStage = gainStageMaps[currentFrame - 1].data[id];
@@ -40,7 +39,6 @@ template <typename Config, typename AccConfig> struct ClusterFinderKernel {
         if (indexQualifiesAsClusterCenter<Config>(id)) {
           findClusterSumAndMax<Config>(energyMaps[currentFrame - 1].data, id,
                                        sum, max);
-
           // check cluster conditions
           if ((energy > c * stddev || sum > n * c * stddev) && id == max) {
             auto &cluster = getClusterBuffer(acc, clusterArray, numClusters);
@@ -51,6 +49,7 @@ template <typename Config, typename AccConfig> struct ClusterFinderKernel {
                    !pedestalFallback) {
 
 
+            //! @todo: remove debugging code
             if(id == 0)
               printf("asdfasdfasdf\n");
 
@@ -65,6 +64,6 @@ template <typename Config, typename AccConfig> struct ClusterFinderKernel {
         processInput(acc, detectorData[currentFrame], gainMaps, pedestalMaps,
                      initPedestalMaps, gainStageMaps[currentFrame],
                      energyMaps[currentFrame], mask, id, pedestalFallback);
-      //}
+    }
   }
 };

### 2020-02-24 src/bench.cpp

@@ -33,7 +33,7 @@ constexpr auto framesPerStageG2 = Values<std::size_t, 999>();
 constexpr auto dimX = Values<std::size_t, 1024>();
 constexpr auto dimY = Values<std::size_t, 512>();
 constexpr auto sumFrames = Values<std::size_t, 2>(); //, 10, 20, 100>();
-constexpr auto devFrames = Values<std::size_t, 10>(); //, 100>(); //, 1000>();
+constexpr auto devFrames = Values<std::size_t, 1>(); //10, 100>(); //, 1000>();
 constexpr auto movingStatWindowSize = Values<std::size_t, 100>();
 constexpr auto clusterSize = Values<std::size_t, 2>(); //, 3, 7, 11>();
 constexpr auto cs = Values<std::size_t, 5>();
@@ -96,11 +96,86 @@ std::vector<Duration> benchmark(unsigned int iterations,
         if (benchmarkingInput.clusters)
             benchmarkingInput.clusters->used = 0;
         auto dispenser = calibrate(benchmarkingInput);
+
+
+        /*
+        //! @todo: remove all this debugging code
+          // download and save std dev of the initial pedestal map                                                                                                                                                                                                   
+  auto initPed = dispenser.downloadInitialPedestaldata();
+  std::ofstream outPede("pede_stddev.bin", std::ios::binary);
+  InitPedestal *pedePtr = alpakaNativePtr(initPed.data)->data;
+  for (unsigned int y = 0; y < Config::DIMY; ++y)
+    for (unsigned int x = 0; x < Config::DIMX; ++x)
+      outPede.write(reinterpret_cast<char *>(&(pedePtr[y * 1024 + x].stddev)),
+                    sizeof(double));
+  outPede.flush();
+  outPede.close();
+
+  outPede.open("pede_m.bin", std::ios::binary);
+  pedePtr = alpakaNativePtr(initPed.data)->data;
+  for (unsigned int y = 0; y < Config::DIMY; ++y)
+    for (unsigned int x = 0; x < Config::DIMX; ++x)
+      outPede.write(
+          reinterpret_cast<char *>(&(pedePtr[y * Config::DIMX + x].m)),
+          sizeof(pedePtr[y * Config::DIMX + x].m));
+  outPede.flush();
+  outPede.close();
+
+  outPede.open("pede_m2.bin", std::ios::binary);
+  pedePtr = alpakaNativePtr(initPed.data)->data;
+  for (unsigned int y = 0; y < Config::DIMY; ++y)
[…]

### 2020-02-24 src/main.cpp

@@ -13,7 +13,7 @@
 #ifdef ALPAKA_ACC_GPU_CUDA_ENABLED
 template <std::size_t TMapSize> using Accelerator = GpuCudaRt<TMapSize>;
 #else
-template <std::size_t TMapSize> using Accelerator = CpuOmp2Blocks<TMapSize>;
+template <std::size_t TMapSize> using Accelerator = CpuSerial<TMapSize>;
 #endif
 // CpuOmp2Blocks<MAPSIZE>;
 // CpuTbbRt<MAPSIZE>;
@@ -48,7 +48,7 @@ const std::string pedestalPath =
     "../../../data_pool/px_101016/allpede_250us_1243__B_000000.dat";
 const std::string gainPath = "../../../data_pool/px_101016/gainMaps_M022.bin";
 const std::string dataPath =
-  "../../../data_pool/px_101016/Insu100.dat";
+  "../../../data_pool/px_101016/Insu_6_tr_1_45d_250us__B_000000_3.dat";
 //"../../../data_pool/px_101016/Insu_6_tr_1_45d_250us__B_000000.dat";
 #endif
 
@@ -98,7 +98,7 @@ auto main(int argc, char* argv[]) -> int
 #endif
 
     // upload and calculate pedestal data
-    dispenser.uploadPedestaldata(pedestaldata);
+   dispenser.uploadPedestaldata(pedestaldata);
 
     // allocate space for output data
     FramePackage<typename Config::EnergyMap, ConcreteAcc> energy_data(
@@ -140,9 +140,9 @@ auto main(int argc, char* argv[]) -> int
 
     // set execution flags
     ExecutionFlags ef;
-    ef.mode = 3;
+    ef.mode = 0;
     ef.summation = 0;
-    ef.masking = 0;
+    ef.masking = 1;
     ef.maxValue = 0;
 
     using EnergyPackageView =
@@ -154,6 +154,49 @@ auto main(int argc, char* argv[]) -> int
     using MaxValuePackageView = FramePackageView_t<EnergyValue, ConcreteAcc>;
 
 
+
+
+    save_image<Config>("mask",
+                       dispenser.downloadMask(), 0);
+
+
+    save_image<Config>(
+        "pedestal", alpakaNativePtr(dispenser.downloadPedestaldata().data), 0);
+    
+  // download and save std dev of the initial pedestal map                                                                                                                                                                                                   
[…]

### 2020-02-24 include/bench.hpp
Fixed bug by implementing one kernel with element layer and one without.

@@ -171,21 +171,6 @@ auto bench(
   typename Config::template ClusterArray<ConcreteAcc> *clusters =
       benchmarkingConfig.clusters;
 
-
-
-
-
-  //! @todo: remove debugging code
-  // write out initial pedestal data
-  save_image<Config>(
-                     "pedestal_start",
-                     alpakaNativePtr(dispenser.downloadPedestaldata().data), 0);
-
-
-
-
-
-  
   // process data maps
   while (offset < benchmarkingConfig.data.numFrames) {
 
@@ -217,31 +202,6 @@ auto bench(
                                            benchmarkingConfig.ef, energy,
                                            photons, sum, maxValues, clusters));
 
-
-
-
-
-
-
-
-    //! @todo: remove debugging code
-    // write out pedestal and image values
-            std::get<1>(futures.back()).wait();
-        dispenser.synchronize();
-
-        save_image<Config>("pedestal_" + std::to_string(offset),
-            alpakaNativePtr(dispenser.downloadPedestaldata().data), 0);
-
-        if (energy)
-          save_image<Config>("energy_" + std::to_string(offset),
-                             alpakaNativePtr(energy->data), 0);
-
-
-
-
-
-
-        
     auto offset_diff = std::get<0>(*futures.rbegin()) - offset;
     offset = std::get<0>(*futures.rbegin());
     sum_offset += (offset_diff + Config::SUM_FRAMES - 1) / Config::SUM_FRAMES;

### 2020-02-24 include/check.hpp

@@ -61,11 +61,7 @@ bool checkResult(tl::optional<T> result, std::string referencePath) {
   if (readBytes < dataSize || !reference_file.read(&dummyByte, 1).eof()) {
     std::cerr << "Number of read bytes does not match (" << readBytes << " vs. "
               << dataSize << "; eof: " << reference_file.eof() << ")!\n";
-
-
-
-    //! @todo: remove debugging code
-    //return false;
+    return false;
   }
 
   // compare data

### 2020-02-24 include/jungfrau-photoncounter/Dispenser.hpp

@@ -144,53 +144,6 @@ public:
         while (offset <= data.numFrames - TConfig::DEV_FRAMES) {
             offset += calcPedestaldata(alpakaNativePtr(data.data) + offset,
                                        TConfig::DEV_FRAMES);
-
-            
-            /*
-            //! @todo: remove debugging code
-            if(offset >= 990 && offset < 1000) {
-                
-              // download and save std dev of the initial pedestal map                                                                                                                                                                                                   
-              auto initPed = downloadInitialPedestaldata();
-              std::ofstream outPede("pede/stddev_" + std::to_string(offset) + ".txt");
-              InitPedestal *pedePtr = alpakaNativePtr(initPed.data)->data;
-              for (unsigned int y = 0; y < TConfig::DIMY; ++y) {
-                outPede << "\t";
-                for (unsigned int x = 0; x < TConfig::DIMX; ++x) {
-                  outPede << pedePtr[y * TConfig::DIMX + x].stddev << " ";
-                }
-                outPede << "\n";
-              }
-              outPede.flush();
-              outPede.close();
-            
-              outPede.open("pede/m_" + std::to_string(offset) + ".txt");
-              pedePtr = alpakaNativePtr(initPed.data)->data;
-              for (unsigned int y = 0; y < TConfig::DIMY; ++y) {
-                outPede << "\t";
-                for (unsigned int x = 0; x < TConfig::DIMX; ++x) {
-                  outPede << pedePtr[y * TConfig::DIMX + x].m << " ";
-                }
-                outPede << "\n";
-              }
-              outPede.flush();
-              outPede.close();
-            
-              outPede.open("pede/m2_" + std::to_string(offset) + ".txt");
-              pedePtr = alpakaNativePtr(initPed.data)->data;
-              for (unsigned int y = 0; y < TConfig::DIMY; ++y) {
-                outPede << "\t";
-                for (unsigned int x = 0; x < TConfig::DIMX; ++x) {
-                  outPede << pedePtr[y * TConfig::DIMX + x].m2 << " ";
-                }
-                outPede << "\n";
-              }
-              outPede.flush();
-              outPede.close();
-            }
-            */
-
-            
         }
 
         // upload remaining frames

### 2020-02-24 include/jungfrau-photoncounter/kernel/ClusterFinder.hpp

@@ -1,6 +1,72 @@
 #pragma once
 #include "helpers.hpp"
+#include "Config.hpp"
 
+// type definition
+template <typename Config, typename AccConfig> struct ClusterFinderKernel;
+
+// type-trait workaround for CUDA
+#ifdef ALPAKA_ACC_GPU_CUDA_ENABLED
+template <typename Config> struct ClusterFinderKernel<Config, GpuCudaRt<Config::MAPSIZE>> {
+  template <typename TAcc, typename TDetectorData, typename TGainMap,
+            typename TInitPedestalMap, typename TPedestalMap,
+            typename TGainStageMap, typename TEnergyMap, typename TClusterArray,
+            typename TNumClusters, typename TMask, typename TNumFrames,
+            typename TCurrentFrame, typename TNumStdDevs = int>
+  ALPAKA_FN_ACC auto operator()(
+      TAcc const &acc, TDetectorData const *const detectorData,
+      TGainMap const *const gainMaps, TInitPedestalMap *const initPedestalMaps,
+      TPedestalMap *const pedestalMaps, TGainStageMap *const gainStageMaps,
+      TEnergyMap *const energyMaps, TClusterArray *const clusterArray,
+      TNumClusters *const numClusters, TMask *const mask,
+      TNumFrames const numFrames, TCurrentFrame const currentFrame,
+      bool pedestalFallback, TNumStdDevs const c = Config::C) const -> void {
+    auto id = getLinearIdx(acc);
+
+    // check range
+    if (id >= Config::MAPSIZE)
+      return;
+
+    constexpr auto n = Config::CLUSTER_SIZE;
+
+    if (currentFrame) {
+      auto adc = getAdc(detectorData[currentFrame - 1].data[id]);
+
+      const auto &gainStage = gainStageMaps[currentFrame - 1].data[id];
+
+      float sum;
+      decltype(id) max;
+
+      const auto &energy = energyMaps[currentFrame - 1].data[id];
+      const auto &stddev = initPedestalMaps[gainStage][id].stddev;
+
+      if (indexQualifiesAsClusterCenter<Config>(id)) {
+        findClusterSumAndMax<Config>(energyMaps[currentFrame - 1].data, id, sum,
+                                     max);
+
+        // check cluster conditions
+        if ((energy > c * stddev || sum > n * c * stddev) && id == max) {
+          auto &cluster = getClusterBuffer(acc, clusterArray, numClusters);
+          copyCluster<Config>(energyMaps[currentFrame - 1], id, cluster);
+        }
[…]

### 2020-02-24 src/bench.cpp

@@ -33,7 +33,7 @@ constexpr auto framesPerStageG2 = Values<std::size_t, 999>();
 constexpr auto dimX = Values<std::size_t, 1024>();
 constexpr auto dimY = Values<std::size_t, 512>();
 constexpr auto sumFrames = Values<std::size_t, 2>(); //, 10, 20, 100>();
-constexpr auto devFrames = Values<std::size_t, 1>(); //10, 100>(); //, 1000>();
+constexpr auto devFrames = Values<std::size_t, 10>(); //, 100>(); //, 1000>();
 constexpr auto movingStatWindowSize = Values<std::size_t, 100>();
 constexpr auto clusterSize = Values<std::size_t, 2>(); //, 3, 7, 11>();
 constexpr auto cs = Values<std::size_t, 5>();
@@ -64,7 +64,6 @@ template <class Tuple> struct ConfigFrom {
                                   ClusterSize::value,
                                   C::value>;
 };
-
 using Duration = std::chrono::nanoseconds;
 using Timer = std::chrono::high_resolution_clock;
 
@@ -95,87 +94,13 @@ std::vector<Duration> benchmark(unsigned int iterations,
     for (unsigned int i = 0; i < iterations; ++i) {
         if (benchmarkingInput.clusters)
             benchmarkingInput.clusters->used = 0;
-        auto dispenser = calibrate(benchmarkingInput);
-
-
-        /*
-        //! @todo: remove all this debugging code
-          // download and save std dev of the initial pedestal map                                                                                                                                                                                                   
-  auto initPed = dispenser.downloadInitialPedestaldata();
-  std::ofstream outPede("pede_stddev.bin", std::ios::binary);
-  InitPedestal *pedePtr = alpakaNativePtr(initPed.data)->data;
-  for (unsigned int y = 0; y < Config::DIMY; ++y)
-    for (unsigned int x = 0; x < Config::DIMX; ++x)
-      outPede.write(reinterpret_cast<char *>(&(pedePtr[y * 1024 + x].stddev)),
-                    sizeof(double));
-  outPede.flush();
-  outPede.close();
-
-  outPede.open("pede_m.bin", std::ios::binary);
-  pedePtr = alpakaNativePtr(initPed.data)->data;
-  for (unsigned int y = 0; y < Config::DIMY; ++y)
-    for (unsigned int x = 0; x < Config::DIMX; ++x)
-      outPede.write(
[…]

### 2020-02-24 Publication/manuscript.tex
Added more bullet points to paper.

@@
 
-\subsection{Alpaka implementation of fast data processing (Jonas, Florian)}
+\subsection{Alpaka implementation of fast data processing (Jonas)}
 \label{subsec:alpaka}
@@
 	\item real-time data processing (at least 1 detector at 2.5 kHz)
-	\item simple and extensible design(?)
 	\item portable to other architectures (CPU OpenMP, TBB, ...) with minimal code changes
+	\item library for easy integration into existing projects
 \end{itemize}
@@
 \begin{itemize}
-	\item communication speed between host and GPU
+	\item communication speed between host and device
 	\item RAM speed
 \end{itemize}
-\textbf{design overview:}
+features:
 \begin{itemize}
-	\item wrap for most important alpaka functions (for ease of use) --$>$ explain alpaka shortly (?)
-	\item templated for maximal performance and configurability (?)
-	\item written in modern C++ for optimal performance
+	\item automatic pixel masking (wrong gainstage in calibration; stddev threshold; user supplied mask)
+	\item max value extraction
+	\item pedestal updates $\rightarrow$ runtime fallback to initial pedestalmaps also implemented
+\end{itemize}
+\textbf{design overview:}\\
+upload strategies:
+\begin{itemize}
+	\item split frames and upload them $\rightarrow$ very inefficient
+	\item upload blocks sequentially $\rightarrow$ inefficient because application is memory bound
+	\item pipeline copy and kernel operations
+\end{itemize}
+\begin{itemize}
+	\item use alpaka to write platform independent code --$>$ explain alpaka shortly (?)
+	\item written in modern C++ for performance
 	\item pipelined stream processing of data: upload data to GPU -$>$ process -$>$ download again
@@
 \end{itemize}
-\textbf{software architecture:}
+\textbf{software architecture:}\\
 \textit{hardware accelerator concept}
@@
 \end{itemize}
-$<$insert hardware sketch here$>$
+$<$insert hardware sketch here$>$\\
 \textit{the dispenser}
+possible work distribution
 \begin{itemize}
-	\item handles devices (initialization, memory management, etc.)
+	\item thread pool $\rightarrow$ unnecessarily complex
+	\item source-sink relationship $\rightarrow$ every source has a dedicated sink in which data is written $\rightarrow$ enqueueing of source sink pairs
+\end{itemize}
+\begin{itemize}
+	\item handles devices (initialization, memory management, etc.) using a ringbuffer
+	\item usage of double buffering (input data in user provided buffers which is then uploaded to the device, calculated into a new device buffer and downloaded into a user provided output buffer) $\rightarrow$ minimal number of buffers; user is in charge of buffer handling
 	\item enqueues data to next free devices (asynchronously)
@@
 \end{itemize}
-$<$insert dispenser structure here$>$
+$<$insert dispenser structure here$>$\\
 \textit{the kernel structure}
@@
 \end{itemize}
-$<$insert example here$>$
+$<$insert example here$>$\\
 \textit{easy hardware abstraction}
@@
 
-\subsection{Benchmark tests (Jonas, Florian)}
+\subsection{Benchmark tests (Jonas)}
 \label{subsec:benchmark}
 Design, objectives, and evaluation of tests
+
+\textcolor{magenta}{
+\begin{itemize}
+	\item machine specks
+	\item used libraries
+	\item which tests where chosen?
+	\item why were they chosen?
+	\item what are the expectations?
+\end{itemize}
+}
 
@@
 
+\textcolor{magenta}{
+expected:
+\begin{itemize}
+	\item RAM and PCIe bandwidth @ \textasciitilde12 GB/s, after that: multiple nodes necessary
+	\item unbalanced up- and downloads (mainly on GPU)
+	\item multiple different accelerators will be each as fast as the slowest one
+\end{itemize}
+}
+
 Available capacity on GPUs\\
 
+\textcolor{magenta}{
+expected: up to 1000 Frames/s; 100 to 1000 Frames per stream \\
+\textbf{comparison of devices (expected):}
+\begin{itemize}
+	\item after 10 Frames per stream the frame count has diminishing effects on performance (on GPU)
+	\item on CPU: number of Frames per streams doesn't effect performance much
+	\item photon calculation a lot faster (\textasciitilde4x) than energy; clustering in between (this effect will probably be less on CPU)
+	\item number and type of kernels mostly insignificant for GPU; some effects on CPU
+	\item frame summing will speed up almost linearly on less frames on GPU, on CPU not so big differences
+	\item almost linear scaling with number of GPUs
+	\item almost linear scaling with number of cores on CPU
+	\item pedestal updates and masking doesn't effect performance mush
+	\item exception: clustering (mainly on GPU): scales linearly with clusters to download \ calculate
+	\item multiple modules show weak scaling
+	\item effects of cluster size???
+\end{itemize}
+}
+
 Best system
+
+\textcolor{magenta}{
+	expected: 4x V100 System with 100 or 1000 frames per stream @ $>$ 10000 Frames/s
+}
 
@@
 
+\textcolor{magenta}{
+applicable to data from similar sensors (rectangular sensor, consecutive in memory)
+}
+
 calculation on FPGAs in the future
+
+\textcolor{magenta}{
+more backends to come (SYCL, ...)
+}
 

### 2020-03-02 Publication/manuscript.tex
Updated Manuscript.pdf

@@
 	\item almost linear scaling with number of cores on CPU
-	\item pedestal updates and masking doesn't effect performance mush
+	\item pedestal updates and masking doesn't effect performance much
 	\item exception: clustering (mainly on GPU): scales linearly with clusters to download \ calculate
-	\item multiple modules show weak scaling
+	\item multiple modules show weak scaling until bandwidth limit is reached
 	\item effects of cluster size???
@@
 \textcolor{magenta}{
-more backends to come (SYCL, ...)
+more backends to come (SYCL, ...) \\
+AMD GPU support in light of Frontier HPC construction
 }

### 2020-03-02 Publication/manuscript.tex

--- 2020-01-06 Publication/manuscript.tex
@@
 
-{\bf Keywords:} Photon pixel detector, fast data processing, GPU programming, Alpaka \textcolor{magenta}{...}
+{\bf Keywords:} Photon pixel detector, fast data processing, GPU programming, Alpaka...
 
@@
 
-\textcolor{magenta}{
-\textbf{requirements:}
-\begin{itemize}
-	\item real-time data processing (at least 1 detector at 2.5 kHz)
-	\item simple and extensible design(?)
-	\item portable to other architectures (CPU OpenMP, TBB, ...) with minimal code changes
-\end{itemize}
+requirements:
+ - realtime data processing (at least 1 detector at 2.5 kHz)
+ - simple and extensible design
+ - portable to other architectures (CPU OpenMP, TBB, ...) with minimal code changes
+
 optional requirements:
-\begin{itemize}
-	\item easily adaptable to similar detectors (e.g. MÖNCH)
-\end{itemize}
-expected bottlenecks:
-\begin{itemize}
-	\item communication speed between host and GPU
-	\item RAM speed
-\end{itemize}
-\textbf{design overview:}
-\begin{itemize}
-	\item wrap for most important alpaka functions (for ease of use) --$>$ explain alpaka shortly (?)
-	\item templated for maximal performance and configurability (?)
-	\item written in modern C++ for optimal performance
-	\item pipelined stream processing of data: upload data to GPU -$>$ process -$>$ download again
-	\item multiple detectors and multiple GPUs usable in parallel
-\end{itemize}
-$<$insert data flow diagram here$>$ \\
-architecture specific optimizations:
-\begin{itemize}
-	\item contiguous memory vs. grid striding loop
-	\item only upload data if needed
-\end{itemize}
-\textbf{software architecture:}
-\textit{hardware accelerator concept}
-\begin{itemize}
-	\item host
-	\item device
-	\item memory not necessarily shared
-	\item data might need to be copied onto device
-	\item device starts (possibly asynchronous) execution
-	\item result is copied back to host
-	\item common host: CPU
-	\item common devices: GPUs, FPGAs, Multicore CPUs (in general multi- and manycore systems)
-\end{itemize}
-$<$insert hardware sketch here$>$
-\textit{the dispenser}
-\begin{itemize}
-	\item handles devices (initialization, memory management, etc.)
-	\item enqueues data to next free devices (asynchronously)
-	\item copies data back from devices (asynchronously)
-\end{itemize}
-$<$insert dispenser structure here$>$
-\textit{the kernel structure}
-\begin{itemize}
-	\item written in empty struct as static function
-	\item all needed values passed to function, no real access to member variables
-	\item most data passed by pointer, only single numbers passed directly
-	\item kernel has access to device properties (block size, grid size, device specific functions, etc.)
-\end{itemize}
-$<$insert example here$>$
-\textit{easy hardware abstraction}
-\begin{itemize}
-	\item useful settings defined for every possible device class
-	\item device is selected by using statement (similar to a typedef)
-	\item no additional changes needed
-\end{itemize}
-$<$insert example here$>$
-}
+ - easily adaptable to similar detectors (e.g. MÖNCH)
+
+design overview:
+ - wrap for most important alpaka functions (for ease of use)
+ - templated for maximal performance and configurability
+ - stream processing of data: upload data to GPU -> process -> download again
+ - multiple detectors and multiple GPUs usable in parallel
 

### 2020-03-02 include/jungfrau-photoncounter/Alpakaconfig.hpp
Temporarily removed CUDA-specific  clustering kernel and improved the general case. The previous bug could not be reproduced.

@@ -289,6 +289,7 @@ template <std::size_t MAPSIZE> struct GpuCudaRt {
   using HostView = alpaka::mem::view::ViewSubView<DevHost, T, Dim, Size>;
 
   static constexpr std::size_t STREAMS_PER_DEV = 3;
+  //! @note: only one element per thread allowed for CUDA (see issue #66)
   static constexpr Size elementsPerThread = 1u;
   static constexpr Size threadsPerBlock = 256;
   static constexpr Size blocksPerGrid = (MAPSIZE + 255) / 256;
@@ -318,6 +319,7 @@ template <std::size_t MAPSIZE> struct GpuCudaRtSync {
   using HostView = alpaka::mem::view::ViewSubView<DevHost, T, Dim, Size>;
 
   static constexpr std::size_t STREAMS_PER_DEV = 1;
+  //! @note: only one element per thread allowed for CUDA (see issue #66)
   static constexpr Size elementsPerThread = 1u;
   static constexpr Size threadsPerBlock = 256;
   static constexpr Size blocksPerGrid = (MAPSIZE + 255) / 256;

### 2020-03-02 include/jungfrau-photoncounter/kernel/ClusterFinder.hpp

@@ -5,6 +5,10 @@
 // type definition
 template <typename Config, typename AccConfig> struct ClusterFinderKernel;
 
+
+
+
+/*
 // type-trait workaround for CUDA
 #ifdef ALPAKA_ACC_GPU_CUDA_ENABLED
 template <typename Config> struct ClusterFinderKernel<Config, GpuCudaRt<Config::MAPSIZE>> {
@@ -65,8 +69,68 @@ template <typename Config> struct ClusterFinderKernel<Config, GpuCudaRt<Config::
   }
 };
 #endif
+*/
+
 
 // general implementation
+template <typename Config, typename AccConfig> struct ClusterFinderKernel {
+  template <typename TAcc, typename TDetectorData, typename TGainMap,
+            typename TInitPedestalMap, typename TPedestalMap,
+            typename TGainStageMap, typename TEnergyMap, typename TClusterArray,
+            typename TNumClusters, typename TMask, typename TNumFrames,
+            typename TCurrentFrame, typename TNumStdDevs = int>
+  ALPAKA_FN_ACC auto operator()(
+      TAcc const &acc, TDetectorData const *const detectorData,
+      TGainMap const *const gainMaps, TInitPedestalMap *const initPedestalMaps,
+      TPedestalMap *const pedestalMaps, TGainStageMap *const gainStageMaps,
+      TEnergyMap *const energyMaps, TClusterArray *const clusterArray,
+      TNumClusters *const numClusters, TMask *const mask,
+      TNumFrames const numFrames, TCurrentFrame const currentFrame,
+      bool pedestalFallback, TNumStdDevs const c = Config::C) const -> void {
+    const uint64_t globalId = getLinearIdx(acc);
+    constexpr uint64_t elementsPerThread = AccConfig::elementsPerThread;
+
+    constexpr auto n = Config::CLUSTER_SIZE;
+
+    // iterate over all elements in the thread
+    for (auto id = globalId * elementsPerThread;
+         id < (globalId + 1) * elementsPerThread &&
+           id < Config::MAPSIZE; ++id) {
+      
+      if (currentFrame) {
+        auto adc = getAdc(detectorData[currentFrame - 1].data[id]);
+        const char &gainStage = gainStageMaps[currentFrame - 1].data[id];
+        float sum;
+        decltype(id) max;
+        const auto &energy = energyMaps[currentFrame - 1].data[id];
[…]
